\emph{For problems 1--18, use the dataset shown in the table below.  This is a sample of 15 players in Major League Baseball, chosen from the starting lineups of teams in 2019.  The table shows the team, age, position, height, and salary for each player, as well as several statistics from that season.  These include the number of games they played (G), their batting average (AVE) (the proportion of their at-bats for which they got a hit), and their home runs (HR).}
\begin{center}
{\footnotesize
\begin{tabular}{l l c c c c c r}
\textbf{Name} & \textbf{Team} & \textbf{Age} & \textbf{Height} & \textbf{G} & \textbf{AVE} & \textbf{HR} & \textbf{Salary}\\
\hline
& & & & & & & \\
Cedric Mullins & Orioles & 25 & 173 cm & 22 & .094 & 0 & \$557,500\\
Tim Anderson & White Sox & 26 & 185 cm & 123 & .335 & 18 & \$1,400,000\\
Christin Stewart & Tigers & 25 & 183 cm & 104 & .233 & 10 & \$556,400\\
Alex Gordon & Royals & 35 & 185 cm & 150 & .266 & 13 & \$20,000,000\\
Jonathan Schoop & Twins & 27 & 185 cm & 121 & .256 & 23 & \$7,500,000\\
Marcus Semien & Athletics & 29 & 183 cm & 162 & .285 & 33 & \$5,900,000\\
Yandy Diaz & Rays & 28 & 188 cm & 79 & .267 & 14 & \$558,400\\
Randal Grichuk & Blue Jays & 28 & 188 cm & 151 & .232 & 31 & \$5,000,000\\
Josh Donaldson & Braves & 33 & 185 cm & 155 & .259 & 37 & \$23,000,000\\
Joey Votto & Reds & 36 & 188 cm & 142 & .261 & 15 & \$25,000,000\\
Cody Bellinger & Dodgers & 24 & 193 cm & 156 & .305 & 47 & \$605,000\\
Ryan Braun & Brewers & 35 & 188 cm & 144 & .285 & 22 & \$19,000,000\\
Maikel Franco & Phillies & 27 & 185 cm & 123 & .234 & 17 & \$5,200,000\\
Ian Kinsler & Padres & 37 & 183 cm & 87 & .217 & 9 & \$3,750,000\\
Marcell Ozuna & Cardinals & 28 & 185 cm & 130 & .241 & 29 & \$12,250,000\\
\end{tabular}}
\end{center}

\begin{center}
\textbf{Note: for problems 1--18, use calculator to find answer; no work shown}
\end{center}

\begin{enumerate}
\item Find the mean age of the players. \answer{29.53 years}

\item Find the mean height of the players. \answer{185.13 cm}

\item Find the mean salary for the players. \answer{\$8,685,153}

\item Find the median number of games played (G) for the players. \answer{130 games}

\item Find the median number of home runs (HR) for the players. \answer{21.2 home runs}

\item Find the median batting average (AVE) for the players. \answer{.259}

\item Find the mode(s) for the height of the players. \answer{185 cm}

\item Find the mode(s) for the ages of the players. \answer{28 years}

\item For the players' ages, which is greater, the mean or the median? \answer{Mean}

\item For the players' salaries, which is greater, the mean or the median? \answer{Mean}

\item Find the range of the ages of the players. \answer{13 years}

\item Find the range of the heights of the players. \answer{20 cm}

\item Find the range of the number of games played by the players. \answer{140 games}

\item Find the standard deviation for the players' salaries. \answer{\$8,839,865}

\item Find the standard deviation for the number of home runs hit by the players. \answer{12.3 home runs}

\item Find the standard deviation for the players' batting averages. \answer{.053}

\item Calculate the Five Number Summary for the players' heights. \answer{173, 183, 185, 188, 193}

\item Calculate the Five Number Summary for the number of games played by the players. \answer{22, 104, 130, 150, 162}

\item Twenty-five randomly selected students were asked how many movies they had watched the previous week.  The results are shown below.
\begin{center}
\begin{tabular}{c c}
\textbf{Number of Movies} & \textbf{Frequency}\\
\hline
& \\
0 & 5\\
1 & 9\\
2 & 6\\
3 & 4\\
4 & 1
\end{tabular}
\end{center}
Find the mean and median for the number of movies watched per student. \answer{Mean: 1.68; Median: 1}
\begin{center}
Mean: $\dfrac{(0)(5) + (1)(9) + (2)(6) + (3)(4) + (4)(1)}{25} = \dfrac{42}{25} = 1.68$\\
Median: $\dfrac{n+1}{2} = \dfrac{26}{2} = 13$, so it is in the 13th position, where there is a 1.
\end{center}

\item Twenty students were asked how many hours they worked per day.  The table below shows the results.
\begin{center}
\begin{tabular}{c c}
\textbf{Number of Hours} & \textbf{Frequency}\\
\hline
& \\
2 & 3\\
3 & 5\\
4 & 3\\
5 & 6\\
6 & 2\\
7 & 1
\end{tabular}
\end{center}
Find the mean and median for the number of hours worked per student. \answer{Mean: 4.1; Median: 4}
\begin{center}
Mean: $\dfrac{(2)(3) + (3)(5) + (4)(3) + (5)(6) + (6)(2) + (7)(1)}{20} = \dfrac{82}{20} = 4.1$\\
Median: $\dfrac{n+1}{2} = \dfrac{21}{2} = 10.5$, so it is between the 10th and 11th positions; average is $\dfrac{4 + 4}{2} = 4$.
\end{center}

\item Find the weighted average of the student's grades listed below, using the given percentage value for each category.
\begin{center}
\begin{tabular}{l c c}
\textbf{Assignment} & \textbf{Score} & \textbf{Weight}\\
\hline
& & \\
Test 1 & 81 & 30\%\\
Test 2 & 88 & 30\%\\
Homework & 92 & 10\%\\
Quizzes & 89 & 10\%\\
Final Exam & 84 & 20\%
\end{tabular}
\end{center}
\text{} \answer{85.6}
\[(81)(0.3) + (88)(0.3) + (92)(0.1) + (89)(0.1) + (84)(0.2) = 85.6\]

\item Find the weighted average of the student's grades listed below, using the given point value for each category.
\begin{center}
\begin{tabular}{l c c}
\textbf{Assignment} & \textbf{Score} & \textbf{Points}\\
\hline
& & \\
Tests & 86 & 200\\
Projects & 94 & 100\\
Homework & 90 & 50\\
Final Exam & 83 & 150
\end{tabular}
\end{center}
\text{} \answer{87.1}
\[\dfrac{(86)(200) + (94)(100) + (90)(50) + (83)(150)}{200 + 100 + 50 + 150} = 87.1\]

\item In a neighborhood donut shop, one type of donut has 530 calories, three types of donuts have 330 calories, four types of donuts have 320 calories, seven types of donuts have 410 calories, and five types of donuts have 380 calories. Find the mean and median calories of the donuts. \answer{Mean: 378.5; Median: 410}
\begin{center}
Mean: $\dfrac{(1)(530) + (3)(330) + (4)(320) + (7)(410) + (5)(380)}{1 + 3 + 4 + 7 + 5} = \dfrac{7570}{20} = 378.5$\\
Median: $\dfrac{n+1}{2} = \dfrac{21}{2} = 10.5$, so it is between the 10th and 11th positions; average is $\dfrac{410 + 410}{2} = 410$.
\end{center}

\item In a recent issue of \textit{IEEE Spectrum}, 84 engineering conferences were announced. Four conferences lasted 2 days, 36 lasted 3 days, 18 lasted 4 days, 19 lasted 5 days, 4 lasted 6 days, 1 lasted 7 days, 1 lasted 8 days, and 1 lasted 9 days. Find the mean and median length (in days) of an engineering conference. \answer{Mean: 3.94; Median: 4}
\begin{center}
Mean: $\dfrac{(4)(2) + (36)(3) + (18)(4) + (19)(5) + (4)(6) + (1)(7) + (1)(8) + (1)(9)}{4 + 36 + 18 + 19 + 4 + 1 + 1 + 1} = \dfrac{331}{84} = 3.94$\\
Median: $\dfrac{n+1}{2} = \dfrac{85}{2} = 42.5$, so it is between the 42nd and 43rd positions; average is $\dfrac{4 + 4}{2} = 4$.
\end{center}
\end{enumerate}