\emph{For problems 1--2, identify the population and sample.}
\begin{enumerate}
\item A political scientist surveys 28 of the current 106 representatives in a state's congress. Of them, 14 said they were supporting a new education bill, 12 said they were not supporting the bill, and 2 were undecided. \answer{Population: 106 representatives; sample: 28 surveyed}

\item The city of Frederick has 9500 registered voters. There are two candidates for the city council in an upcoming election: Marfani and Rahman. The day before the election, a telephone poll of 350 randomly selected registered voters was conducted. Of them, 112 said they would vote for Marfani, 207 said they would vote for Rahman, and 31 were undecided. \answer{Population: 9500 registered voters; sample: 350 randomly selected}
\end{enumerate} 

\emph{For problems 3--4, decide whether the sampling method described is likely to produce a representative sample, and why or why not.}
\begin{enumerate}
\setcounter{enumi}{2}
\item A marriage counselor is interested in the studying divorce rates, so she gives her clients a survey. \answer{No; clients of a marriage counselor are more likely to get divorced}

\item A fitness center wants to see how much use their treadmills get, so they pick random times during the day and record how many treadmills are in use each time. \answer{Yes; no obvious bias}
\end{enumerate}

\emph{For problems 5--10, determine the type of sampling used in the given scenario.}
\begin{enumerate}
\setcounter{enumi}{4}
\item A high school principal polls 50 freshmen, 50 sophomores, 50 juniors, and 50 seniors regarding policy changes for after school activities. \answer{Stratified}

\item To check their accuracy, the Census Bureau draws a sample of several city blocks and recounts everyone in those blocks. \answer{Cluster}

\item A pollster walks around a busy shopping mall and asks people passing by how often they shop at the mall. \answer{Convenience}

\item Police at a DUI checkpoint stop every tenth car to check whether the driver is sober. \answer{Systematic}

\item A restaurant samples 100 sales from the past week by numbering all their receipts, generating 100 random numbers, and picking the receipts that correspond to those numbers. \answer{Simple random}

\item The provost at a university wants to know how a particular policy is affecting faculty, so she randomly selects 3 members of each department to survey. \answer{Stratified}
\end{enumerate}