\newgeometry{margin=0.75in}
\textbf{This text is licensed under a Creative Commons Attribution-Share Alike 3.0 United States License.}\\

To view a copy of this license, visit \href{http://creativecommons.org/licenses/by-sa/3.0/us/}{http://creativecommons.org/licenses/by-sa/3.0/us/} or send a letter to Creative Commons, 171 Second Street, Suite 300, San Francisco, California, 94105, USA\\

You are \textbf{free}:\\
\hspace*{0.25in} \textbf{to Share} -- to copy, distribute, display, and perform the work\\
\hspace*{0.25in} \textbf{to Remix} -- to make derivative works\\

Under the following conditions:\\
\hspace*{0.25in} \textbf{Attribution.}  You must attribute the work in the manner specified by the author or licensor (but not in any way that suggests that they endorse you or your use of the work).\\
\hspace*{0.25in} \textbf{Share Alike.}  If you alter, transform, or build upon this work, you may distribute the resulting work only under the same, similar, or a compatible license.\\

With the understanding of the following:\\
\hspace*{0.25in} \textbf{Waiver.}  Any of the above conditions can be waived if you get permission from the copyright holder.\\
\hspace*{0.25in} \textbf{Other Rights.}  In no way are any of the following rights affected by the license:
\begin{itemize}
\item Your fair dealing or fair use rights
\item Apart from the remix rights granted under this license, the authors' moral rights
\item Rights other persons may have either in the work itself or in how the work is used, such as publicity or privacy rights
\item Notice --- For any reuse or distribution, you must make clear to others the license terms of this work.  The best way to do this is with a link to the following web page: \href{http://creativecommons.org/licenses/by-sa/3.0/us/}{http://creativecommons.org/licenses/by-sa/3.0/us/}
\end{itemize}

\paragraph{Attributions} This book benefited tremendously from others who went before and freely shared their creative work.  The following is a short list of those whom we have to thank for their work and their generosity in contributing to the free and open sharing of knowledge.
\begin{itemize}
\item David Lippman, author of \textit{Math in Society}.  This book uses sections derived from his chapters on Finance, Growth Models, and Statistics.  He also administers MyOpenMath, the free online homework portal to which the problems in this text were added.
\item The developers of \href{onlinestatbook.com}{onlinestatbook.com}.
\item OpenStax College (their book \textit{Introductory Statistics} was used as a reference)

OpenStax College, \textit{Introductory Statistics}. OpenStax College. 19 September 2013. <http://cnx.org/content/col11562/latest/>
\item The authors of \href{https://www.openintro.org/stat/?stat_book=os}{OpenIntro Statistics}, which was also used as a reference.
\item The Saylor Foundation Statistics Textbook: http://www.saylor.org/site/textbooks/Introductory$\%$20Statistics.pdf
\end{itemize}

\paragraph{Thanks} The following is a short list of those whom we wish to thank for their help and support with this project.
\begin{itemize}
\item The President's office at Frederick Community College, for providing a grant to write the first chapters.
\item Gary Hull, who in his tenure as department chair gave us his full support and gave us the impetus to start the project, and generously shared his notes for MA 103.
\item The entire FCC math department, who provided untold support and encouragement, as well as aid in reviewing and editing the text.
\end{itemize}

%\restoregeometry




















