\begin{exercises}
\ptwo{A jar contains 6 red marbles numbered 1 to 6 and 8 blue marbles numbered 1 to 8. A marble is drawn at random from the jar. Find the probability the marble is red or odd-numbered.}
\ptwo{A jar contains 4 red marbles numbered 1 to 4 and 10 blue marbles numbered 1 to 10. A marble is drawn at random from the jar. Find the probability the marble is blue or even-numbered.}

\ptwo{You draw one card from a standard 52-card deck. 
\begin{enumerate}[(a)]
\item What is the probability of selecting a King or a Queen?
\item What is the probability of selecting a face card or a 10?
\item What is the probability of selecting a spade or a heart?
\item What is the probability of selecting a red card or a black card?
\end{enumerate}}
\ptwo{You are dealt a single card from a standard 52-card deck.
\begin{enumerate}[(a)]
\item Find the probability that you are not dealt a diamond.
\item Find the probability that you are not dealt a face card.
\item Find the probability that you are not dealt an Ace.
\item Find the probability that you are not dealt a jack or a king.
\end{enumerate}}

\pone{Consider the following information about a group of 130 FCC students:
\begin{center}
\begin{tabular}{c | c c | c}
\textbf{Gender} & \textbf{Right-handed} & \textbf{Left-handed} & \textbf{Total} \\ \hline 
\textbf{Female} & 58 & 13 & 71\\
\textbf{Male} & 47 & 12 & 59  \\ \hline
\textbf{Total} & 105 & 25 & 130 
\end{tabular}
\end{center}
If one person is randomly selected from the group, what is the probability this student is female or left-handed?}

\pone{The table below shows the number of survey subjects who have received and not received a speeding ticket in the last year, and the color of their car. Find the probability that a randomly chosen person has a red car \emph{or} got a speeding ticket. 
\begin{center}
\begin{tabular}{c | c c | c}
& \textbf{Speeding Ticket} & \textbf{No Speeding Ticket }& \textbf{Total} \\ \hline 
\textbf{Red Car} & 15 & 135 & 150\\
\textbf{Not Red Car} & 45 & 470 & 515  \\ \hline
\textbf{Total} & 60 & 605 & 665
\end{tabular}
\end{center}}

\ptwo{Suppose you roll a blue six sided die and a red six sided die, and add their totals.
Find the probability of rolling:
\begin{enumerate}[(a)]
\item A 7 or 11.
\item An even number or a number less than 6.
\item A prime number or a number greater than 5.
\end{enumerate}}
\ptwo{A bag contains 4 white counters, 6 black counters, and 1 green counter. What is the
probability of drawing:
\begin{enumerate}[(a)]
\item A white counter or a green counter?
\item A black counter or a green counter?
\item Not a green counter?
\end{enumerate}}

\pone{A poll was taken of 14,056 working adults aged 40-70 to determine their level of
education. The participants were classified by sex and by level of education. The results
were as follows.
\begin{center}
\begin{tabular}{c|cc|c}
\textbf{Education Level}     &  \textbf{Male} &  \textbf{Female} &  \textbf{Total} \\ \hline
\textbf{High School or Less} & 3141  &2434     &  5575  \\ 
\textbf{Bachelor's Degree}   & 3619  &3761     &  7380  \\
\textbf{Master's Degree}     & 534   &472      &  1006  \\
\textbf{Ph.D.}               & 52    &43       &  95    \\ \hline
\textbf{Total} 							& 7346  &6710     & 14,056 \\ 
\end{tabular}
\end{center}
A person is selected at random. Compute the following probabilities:
\begin{enumerate}[(a)]
\item The probability that the selected person does not have a Ph.D.
\item The probability that the selected person does not have a Master's degree.
\item The probability that the selected person is female or has a Master's degree.
\item The probability that the selected person is male or has a Ph.D.
\end{enumerate}}
\end{exercises}