\begin{exercises}
\pone{You have a box of chocolates that contains 50 pieces, of which 30 are solid chocolate,
15 are filled with cashews and 5 are filled with cherries. All the candies look exactly alike.
You select a piece, eat it, select a second piece, eat it, and finally eat one last piece. Find the probability of selecting a solid chocolate followed by two cherry-filled chocolates.}

\pone{You roll a fair six-sided die twice. Find the probability of rolling a 6 the first time
and a number greater than 2 the second time.}

\ptwo{A math class consists of 25 students, 14 female and 11 male. Three students are selected at random, one at a time, to participate in a probability experiment. Compute the probability that:
\begin{enumerate}[(a)]
\item A male is selected, then two females.
\item A female is selected, then two males.
\item Two females are selected, then one male.
\item Three males are selected.
\item Three females are selected.
\end{enumerate}
}
\ptwo{A large cooler contains the following drinks: 6 lemonade, 8 Sprite, 15 Coke, and 7 root beer. You randomly pick two cans, one at a time (without replacement).
\begin{enumerate}[(a)]
\item What is the probability that you get 2 cans of Sprite?
\item What is the probability that you do not get 2 cans of Coke?
\item What is the probability that you get either 2 root beer or 2 lemonade?
\item What is the probability that you get one can of Coke and one can of Sprite?
\item What is the probability that you get two drinks of the same type?
\end{enumerate}}

\pone{My top drawer contains different colored socks: 14 are white, 10 are black, 6 are pink, and 4 are blue. All socks in the drawer are loose. Every morning I randomly select 2 socks, one at a time. Calculate the following probabilities, giving both fraction and decimal answers, rounding to 4 decimal places:
\begin{enumerate}[(a)]
\item What is the probability that I get a blue pair of socks?
\item What is the probability that I do not get a blue pair of socks?
\item What is the probability that I either get a white pair or a blue pair of socks?
\item What is the probability that I get one black sock and one white sock?
\end{enumerate}}

\ptwo{Suppose a math class contains 25 students, 14 females (three of whom speak French) and 11 males (two of whom speak French). Compute the probability that a randomly selected student speaks French, given that the student is female.}
\ptwo{Suppose a math class contains 30 students, 18 females (four of whom speak French) and 12 males (three of whom speak French). Compute the probability that a randomly selected student is male, given that the student speaks French.}

\pone{A certain virus infects one in every 400 people. A test used to detect the virus in a person is positive 90\% of the time if the person has the virus and 10\% of the time if the person does not have the virus.% Let A be the event ``the person is infected'' and B be the event ``the person tests positive.''
\begin{enumerate}[(a)]
\item Find the probability that a person has the virus given that they have tested positive.
\item Find the probability that a person does not have the virus given that they test negative. 
\end{enumerate}}

\pone{A poll was taken of 14,056 working adults aged 40-70 to determine their level of
education. The participants were classified by sex and by level of education. The results
were as follows.
\begin{center}
\begin{tabular}{c|cc|c}
\textbf{Education Level}     &  \textbf{Male} &  \textbf{Female} &  \textbf{Total} \\ \hline
\textbf{High School or Less} & 3141  &2434     &  5575  \\ 
\textbf{Bachelor's Degree}   & 3619  &3761     &  7380  \\
\textbf{Master's Degree}     & 534   &472      &  1006  \\
\textbf{Ph.D.}               & 52    &43       &  95    \\ \hline
\textbf{Total} 							& 7346  &6710     & 14,056 \\ 
\end{tabular}
\end{center}
A person is selected at random. Compute the following probabilities:
\begin{enumerate}
\item The probability that the selected person is male, given he has a Master's degree.
\item The probability that the selected person does not have  a Master's degree, given it is a male.
\item The probability that the selected person is female, given that she has a Bachelor's degree.
\item The probability that the selected person has a Ph.D, given it is a female.
\end{enumerate}}
\end{exercises}