\begin{exercises}
\textit{For problems 1--8, convert each number to a percent.}\\
\pfour{$\dfrac{1}{2}$}
\pfour{$0.04$}
\pfour{$\dfrac{3}{5}$}
\pfour{$0.79$}

\pfour{1.35}
\pfour{$\dfrac{10}{4}$}
\pfour{12.5}
\pfour{0.0378}

\textit{For problems 9--16, convert each percent to a decimal.}\\
\pfour{33\%}
\pfour{2.6\%}
\pfour{124\%}
\pfour{1240.5\%}

\pfour{42\%}
\pfour{4.5\%}
\pfour{0.003\%}
\pfour{$\dfrac{1}{4}\%$}

\pthree{What is 12\% of 72?}
\pthree{Ninety is what percentage of 200?}
\pthree{Twenty is 15\% of what?}

\pthree{Six is what percentage of 40?}
\pthree{Forty is 91\% of what?}
\pthree{What is 65\% of 65?}

\ptwo{In the fall of 2009 FCC enrolled 6,233 students.  Of those enrolled, 2,810 are in the 18-21 age group.  What percent of FCC students does this represent?}
\ptwo{Patrick left an \$8 tip on a \$50 restaurant bill.  What percent tip is that?}

\ptwo{Ireland has a 23\% VAT (value-added tax, similar to a sales tax).  How much will the VAT be on a purchase of a \euro 250 item?}
\ptwo{Employees in 2012 paid 4.2\% of their gross wages towards social security (FICA tax).  How much would someone earning \$45,000 a year pay toward social security?}

\ptwo{A project on Kickstarter was aiming to raise \$15,000 for a precision coffee press.  They ended up with 714 supporters, raising 557\% of their goal.  How much did they raise?}
\ptwo{Another Kickstarter project for an iPad stylus raised 1,253\% of their goal, finishing with a total of \$313,250 from 7,511 supporters.  What was their original goal?}

\ptwo{One year ago the median price for a home was \$275,000.  Now the current median price for a home is \$235,000.  What was the percent decrease in the median price of a home over the last year?}
\ptwo{There were 943 tornadoes reported in the U.S. in 2013, and 897 tornadoes were reported in 2014.  What percent decrease was there from 2013 to 2014?}

\ptwo{The population of a town increased from 3,250 in 2008 to 4,300 in 2010.  Find the absolute and percent increase.}
\ptwo{The number of CDs sold in 2010 was 114 million, down from 147 million the previous year.  Find the absolute and percent decrease.}

\ptwo{A company wants to decrease their energy use by 15\%.
\begin{enumerate}[(a)]
\item If their electric bill is currently \$2,200 a month, what will their bill be if they're successful?
\item If their next bill is \$1,700 a month, were they successful?  What percent decrease was there from the current bill?
\end{enumerate}
}
\ptwo{A store is hoping an advertising campaign will increase their number of customers by 30\%.  They currently have about 80 customers per day.
\begin{enumerate}[(a)]
\item How many customers will they have if their campaign is successful?
\item If they increase to 120 customers a day, were they successful?  What percent increase is this from the current level?
\end{enumerate}
}

\ptwo{An article reports that ``attendance dropped 6\% this year, to 300.''  What was the attendance before the drop?}
\ptwo{An article reports that ``sales have grown by 30\% this year, to \$200 million.''  What were sales before the growth?}

\ptwo{The U.S. federal debt at the end of 2001 was \$5.77 trillion, and grew to \$6.20 trillion by the end of 2002. At the end of 2005 it was \$7.91 trillion, and grew to \$8.45 trillion by the end of 2006. Calculate the absolute and relative increase for 2001-2002 and 2005-2006. Which year saw a larger relative increase in federal debt?}
\ptwo{A TV originally priced at \$799 is on sale for 30\% off. There is then a 9.2\% sales tax. Find the price after including the discount and sales tax.}

\ptwo{The Walden University had 47,456 students in 2010, while Kaplan University had 77,966 students.  Complete the following statements.
\begin{enumerate}[(a)]
\item Kaplan's enrollment was \line(1,0){20} \% larger than Walden's.
\item Walden's enrollment was \line(1,0){20} \% smaller than Kaplan's.
\item Walden's enrollment was \line(1,0){20} \% of Kaplan's.
\end{enumerate}}
\ptwo{In the 2012 Olympics, Usain Bolt ran the 100 m dash in 9.63 seconds.  Jim Hines won the 1968 gold with a time of 9.95 seconds.
\begin{enumerate}[(a)]
\item Bolt's time was \line(1,0){20} \% faster than Hines'.
\item Hines' time was \line(1,0){20} \% slower than Bolt's.
\item Hines' time was \line(1,0){20} \% of Bolt's.
\end{enumerate}}

\ptwo{A store has clearance items that have been marked down by 60\%.  They are having a sale, advertising an additional 30\% off clearance items.  What percent of the original price do you end up paying?}
\ptwo{A publisher marks up a textbook by 65\%, and a bookstore further marks up the textbook by 35\%.  What percentage of the original cost do you pay?}
\end{exercises}