\begin{enumerate}
\item A survey asked 200 people what beverage they drink in the morning, and offered two possible choices: tea and coffee.  Suppose that 20 answered tea only, 80 answered coffee only, and 40 answered both.
\begin{enumerate}[(a)]
\item How many people drink tea in the morning? \answersub{60}
\item How many people drink neither tea nor coffee? \answersub{60}
\end{enumerate}

\item A survey asked 100 people whether they used Twitter or Facebook in the last month.  Of those surveyed, a total of 40 used Twitter, 70 used Facebook, and 20 used both.
\begin{enumerate}[(a)]
\item How many people used only Facebook? \answersub{50}
\item How many people used neither Facebook nor Twitter? \answersub{10}
\end{enumerate}

\item Out of 100 customers of Domino's Pizza, 60 ordered pizza with onions and pepperoni, 80 ordered it with pepperoni, and 72 ordered it with onions.
\begin{enumerate}[(a)]
\item How many ordered onions but not pepperoni? \answersub{12}
\item How many ordered pepperoni but not onions? \answersub{20}
\item How many ordered neither onions nor pepperoni? \answersub{8}
\end{enumerate}

\item Out of 100 students surveyed, 24 rent movies, 20 rent movies and go the theater, and 15 do neither.
\begin{enumerate}[(a)]
\item How many students only rent movies? \answersub{4}
\item How many students only go to the theater? \answersub{61}
\item How many students go to the theater or rent movies? \answersub{85}
\end{enumerate}

\item An independent survey agency was hired by the Metro to find out how many people commute to their school or job.  The agency interviewed 1000 commuters and submitted the following report:
\begin{center}
\begin{tabular}{l l}
631 came by car & 373: car and bus\\
554 came by bus & 301: bus and metro\\
759 came by metro & 268: car and metro\\
\multicolumn{2}{c}{231: all three types of transportation}
\end{tabular}
\end{center}
The Metro refused to accept the report, stating that it was inaccurate.  Why? \answer{The numbers are not consistent with 1000 surveyed (adds up 1233)}

\item One hundred fifty people were surveyed and asked if they believed in UFOs, ghosts, and Bigfoot.
\begin{center}
\begin{tabular}{l l}
43 believed in UFOs & 8: ghosts and Bigfoot\\
25 believed in Bigfoot & 10: UFOs and ghosts\\
44 believed in ghosts & 5: UFOs and Bigfoot\\
\multicolumn{2}{c}{2 believed in all three}
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item How many people surveyed believed in at least one of these things? \answersub{91}
\item How many people believed in ghosts and Bigfoot, but not UFOs? \answersub{6}
\item How many people didn't believe in any of the three? \answersub{59}
\item How many people believed in Bigfoot only? \answersub{14}
\end{enumerate}

\item A survey asked students whether they had seen \textit{Star Wars}, \textit{The Matrix}, or \textit{Lord of the Rings (LotR)}.
\begin{center}
\begin{tabular}{l l}
24 had seen \textit{Star Wars} & 10: \textit{Star Wars} and \textit{The Matrix}\\
18 had seen \textit{The Matrix} & 12: \textit{The Matrix} and \textit{LotR}\\
20 had seen \textit{LotR} & 14: \textit{Star Wars} and \textit{LotR}\\
\multicolumn{2}{c}{6 had seen all three}
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item How many students have seen exactly one of these movies? \answersub{8}
\item How many students have seen only \textit{Star Wars}? \answersub{6}
\item How many students have seen \textit{Star Wars}, but not \textit{LotR}? \answersub{12}
\item How many students have not seen \textit{The Matrix}? \answersub{14}
\end{enumerate}

\item A survey was given asking whether respondents watch movies at home from Netflix, Redbox, or Amazon Video.
\begin{center}
\begin{tabular}{l l}
53 only use Netflix & 48: only Netflix and Redbox\\
62 only use Redbox & 16: only Redbox and Amazon\\
24 only use Amazon Video & 30: only Netflix and Amazon\\
10 use all three & 25: none of these
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item How many people use Redbox? \answersub{136}
\item How many people use at least one of these? \answersub{243}
\item How many people were surveyed? \answersub{268}
\end{enumerate}

\item A survey asked buyers whether color, size, or brand influenced their choice of cell phone.  The results are below.
\begin{center}
\begin{tabular}{l l}
5 only said color & 20: only color and size\\
8 only said size & 53: only size and brand\\
16 only said brand & 42: only color and brand\\
102 said all three & 20: none of these
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item How many people were influenced by brand? \answersub{213}
\item How many people were influenced by color or size? \answersub{230}
\item How many people were surveyed? \answersub{266}
\end{enumerate}
\end{enumerate}