\emph{In problems 1--4, determine whether each argument is inductive or deductive.}
\begin{enumerate}

\item The last mayor was honest.  The current mayor is honest.  All mayors are honest. \answer{Inductive}

\item Every word has the letter \emph{e} in it.  Your name has the letter \emph{e} in it. \answer{Deductive}

\item All deserts have some plant life.  Some plants live in the Gobi Desert. \answer{Deductive}

\item The sun rose yesterday, and it has risen every other day of my life.  Therefore, the sun rises every day. \answer{Inductive}
\end{enumerate}

\emph{In problems 5--6, use a diagram (an Euler diagram) to evaluate whether each argument is valid.}
\begin{enumerate}
\setcounter{enumi}{4}

\item Every concert pianist can play \emph{Chopsticks}.  Ellen can play \emph{Chopsticks}, so she must be a concert pianist. \answer{Invalid}

\item Every computer program has some bugs in it.  Microsoft Excel is a computer program, so it has some bugs. \answer{Valid}
\end{enumerate}
\pagebreak

\emph{In problems 7--14, use a truth table to determine whether the given argument is valid.}
\begin{enumerate}
\setcounter{enumi}{6}

\item 
\begin{tabular}{l}
$p \to \sim q$\\
$q$\\
\hline
$\therefore\ \sim p$
\end{tabular} \text{} \answer{Valid}

\item 
\begin{tabular}{l}
$p \to q$\\
$\sim p$\\
\hline
$\therefore\ q$
\end{tabular} \text{} \answer{Invalid}


\item 
\begin{tabular}{l}
$p \to q$\\
$q \to p$\\
\hline
$\therefore\ p \wedge q$
\end{tabular} \text{} \answer{Invalid}


\item 
\begin{tabular}{l}
$p \to q$\\
$q \to r$\\
\hline
$\therefore\ \sim p \to \sim r$
\end{tabular} \text{} \answer{Invalid}


\item 
\begin{tabular}{l}
$p \to q$\\
$q \wedge r$\\
\hline
$\therefore\ p \vee r$
\end{tabular} \text{} \answer{Valid}


\item 
\begin{tabular}{l}
$p \leftrightarrow q$\\
$q \to r$\\
\hline
$\therefore\ \sim r \to \sim p$
\end{tabular} \text{} \answer{Valid}


\item 
\begin{tabular}{l}
$q \to \sim p$\\
$q \wedge r$\\
\hline
$\therefore\ r \to p$
\end{tabular} \text{} \answer{Invalid}


\item 
\begin{tabular}{l}
$\sim p \wedge q$\\
$p \leftrightarrow r$\\
\hline
$\therefore\ p \wedge r$
\end{tabular} \text{} \answer{Invalid}

\end{enumerate}

\emph{In problems 15--22, determine which of the standard argument forms matches the given argument, and indicate whether this is a valid argument.}
\begin{enumerate}
\setcounter{enumi}{14}

\item If it is cold, my windows frost over.  My windows are not frosted over, so it is not cold. \answer{Contrapositive Reasoning (Valid)}

\item You must eat well or you will not be healthy.  I eat well, therefore I am healthy. \answer{Disjunctive Misuse (Invalid)}

\item We must build a hydroelectric plant or a nuclear plant.  We won't build a nuclear plant, so we must build a hydroelectric plant. \answer{Disjunctive Reasoning (Valid)}

\item If we open the window, we will hear the birds.  We hear the birds.  Therefore, we opened the window. \answer{Fallacy of the Converse (Invalid)}

\item If I'm tired, I'm cranky.  I'm tired.  Therefore, I'm cranky. \answer{Direct Reasoning (Valid)}

\item If everyone obeyed the law, no jails would be needed.  Not everyone obeys the law, so some jails are needed. \answer{Fallacy of the Inverse (Invalid)}

\item If you pass this course, you will graduate.  If you graduate, you will get a job.  Therefore, if you get a job, you must have passed this course. \answer{Transitive Misuse (Invalid)}

\item If you study the old masters, your art will improve.  If your art improves, you will get accepted to the art institute.  Therefore, if you study the old masters, you will get accepted to the art institute. \answer{Transitive Reasoning (Valid)}
\end{enumerate}