\emph{In problems 1--8, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ You studied.}\\
&q: \textrm{ You passed this course.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\begin{enumerate}
\item If you passed this course, you studied. \answer{$q \to p$}

\item You passed this course if and only if you studied. \answer{$q \leftrightarrow p$}

\item You didn't study if and only if you passed this course. \answer{$\sim p \leftrightarrow q$}

\item You didn't study if you didn't pass this course. \answer{$\sim q \to \sim p$}
\end{enumerate}

\emph{Write each of the following statements in words.}
\begin{enumerate}
\setcounter{enumi}{4}

\item $p \to q$ \answer{If you studied, then you passed this course.}

\item $\sim p \leftrightarrow\ \sim q$ \answer{You didn't study if and only if you didn't pass this course.}

\item $\sim q \to p$ \answer{If you didn't pass this course, then you studied.}

\item $\sim p \to q$ \answer{If you didn't study, then you passed this course.}
\end{enumerate}
\pagebreak

\emph{In problems 9--16, let $p$, $q$, and $r$ represent the following statements:}
\begin{align*}
&p: \textrm{ The ice cream truck is here.}\\
&q: \textrm{ The pool is open.}\\
&r: \textrm{ It is summertime.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}
\begin{enumerate}
\setcounter{enumi}{8}

\item If the ice cream truck is here, and the pool is open, then it is summertime. \answer{$(p \wedge q) \to r$}

\item It isn't summertime if the ice cream truck isn't here and the pool isn't open. \answer{$(\sim p \wedge \sim q) \to \sim r$}

\item It's summertime if and only if the pool is open. \answer{$r \leftrightarrow q$}

\item If it isn't summertime, the pool isn't open or the ice cream truck isn't here. \answer{$\sim r \to (\sim q \vee \sim p)$}
\end{enumerate}

\emph{Write each of the following statements in words.}
\begin{enumerate}
\setcounter{enumi}{12}

\item $(p \vee q) \to r$ \answer{If the ice cream truck is here or the pool is open, then it is summertime.}

\item $r \to (p \wedge q)$ \answer{If it is summertime, then the ice cream truck is here and the pool is open.}

\item $p \leftrightarrow q$ \answer{The ice cream truck is here if and only if the pool is open.}

\item $p \leftrightarrow (q \wedge r)$ \answer{The ice cream truck is here if and only if the pool is open and it is summertime.}
\end{enumerate}

\emph{In problems 17--24, fill in the blanks in each truth table.}
\begin{enumerate}
\setcounter{enumi}{16}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c|}
\hline
$p$ & $q$ & $q \to p$ & $\sim (q \to p)$\\
\hline
& & &\\
T & T & T & F\\
T & F & T & F\\
F & T & F & T\\
F & F & T & F\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $\sim q$ & $\sim p \leftrightarrow \sim q$\\
\hline
& & & &\\
T & T & F & F & T\\
T & F & F & T & F\\
F & T & T & F & F\\
F & F & T & T & T\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c|}
\hline
$p$ & $q$ & $p \vee q$ & $q \to (p \vee q)$\\
\hline
& & &\\
T & T & T & T\\
T & F & T & T\\
F & T & T & T\\
F & F & F & T\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $p \wedge q$ & $\sim p$ & $\sim q$ & $\sim p \wedge \sim q$ & $(p \wedge q) \to (\sim p \wedge \sim q)$\\
\hline
& & & & & &\\
T & T & T & F & F & F & F\\
T & F & F & F & T & F & T\\
F & T & F & T & F & F & T\\
F & F & F & T & T & T & T\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $q \to p$ & $p \to q$ & $(q \to p) \wedge (p \to q)$\\
\hline
& & & &\\
T & T & T & T & T\\
T & F & T & F & F\\
F & T & F & T & F\\
F & F & T & T & T\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $p \to q$ & $\sim q$ & $(p \to q) \wedge \sim q$\\
\hline
& & & &\\
T & T & T & F & F\\
T & F & F & T & F\\
F & T & T & F & F\\
F & F & T & T & T\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c | c|}
\hline
$p$ & $q$ & $r$ & $\sim p$ & $\sim p \vee r$ & $(\sim p \vee r) \to q$\\
\hline
& & & & &\\
T & T & T & F & T & T\\
T & F & T & F & T & F\\
F & T & T & T & T & T\\
F & F & T & T & T & F\\
T & T & F & F & F & T\\
T & F & F & F & F & T\\
F & T & F & T & T & T\\
F & F & F & T & T & F\\
\hline
\end{tabular}
\end{center}

\item 
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $r$ & $\sim p$ & $\sim r$ & $q \to \sim p$ & $\sim r \wedge (q \to \sim p)$\\
\hline
& & & & & &\\
T & T & T & F & F & F & F\\
T & F & T & F & F & T & F\\
F & T & T & T & F & T & F\\
F & F & T & T & F & T & F\\
T & T & F & F & T & F & F\\
T & F & F & F & T & T & T\\
F & T & F & T & T & T & T\\
F & F & F & T & T & T & T\\
\hline
\end{tabular}
\end{center}
\end{enumerate}

\emph{In problems 25--28, determine if each statement is a tautology or not.}
\begin{enumerate}
\setcounter{enumi}{24}

\item $p \wedge \sim p$ \answer{No, this is not a tautology.}

\item $p \vee \sim p$ \answer{Yes, this is a tautology.}

\item $(p \wedge q) \vee (\sim p \wedge q) \leftrightarrow q$ \answer{Yes, this is a tautology.}

\item $(q \to p) \vee (\sim q \to \sim p)$ \answer{Yes, this is a tautology.}
\end{enumerate}
\pagebreak

\emph{In problems 29--30, select the statement that is equivalent to the one given.}
\begin{enumerate}
\setcounter{enumi}{28}

\item ``Either the Cardinals or the Yankees will win the World Series.'' \answer{(a) If the Cardinals don't win the World Series, the Yankees will win it.}
\begin{enumerate}[(a)]
\item If the Cardinals don't win the World Series, the Yankees will win it.
\item The Cardinals and the Yankees will win the World Series.
\item If the Cardinals win the World Series, the Yankees will not win it.
\item If the Yankees win the World Series, the Cardinals will not.
\end{enumerate}

\item ``If the light is on, someone is home.'' \answer{(d) If someone is not home, the light is not on.}
\begin{enumerate}[(a)]
\item The light is on and someone is home.
\item If someone is home, the light is on.
\item Either someone is not home, or the light is on.
\item If someone is not home, the light is not on.
\end{enumerate}
\end{enumerate}

\emph{In problems 31--32, write the converse, inverse, and contrapositive of the given statement.}
\begin{enumerate}
\setcounter{enumi}{30}

\item If you drive through the field, you see fireflies.
\begin{itemize}
\item Converse: \boxtext{If you see fireflies, you drive through the field.}
\item Inverse: \boxtext{If you don't drive through the field, you don't see fireflies.}
\item Contrapositive: \boxtext{If you don't see fireflies, you don't drive through the field.}
\end{itemize}

\item If you go to office hours, you pass the test.
\begin{itemize}
\item Converse: \boxtext{If you pass the test, you go to office hours.}
\item Inverse: \boxtext{If you don't go to office hours, you don't pass the test.}
\item Contrapositive: \boxtext{If you don't pass the test, you don't go to office hours.}
\end{itemize}
\end{enumerate}