\emph{In problems 1--8, write the negation of each conditional statement.}
\begin{enumerate}
\item If a fruit is blue, then it is not a banana. \\ 
\begin{framed}
A fruit is blue and it is a banana.
\end{framed}

\item If the storm comes through, that awning will blow away. \\ 
\begin{framed}
The storm comes through and that awning does not blow away.
\end{framed}

\item You'll catch a cold if you don't take Vitamin C. \\ 
\begin{framed}
You don't take Vitamin C and you don't catch a cold.
\end{framed}
\pagebreak

\item You'll get a three percent return on your investment if you invest with us. \\ 
\begin{framed}
You invest with us and you do not get a three percent return on your investment.
\end{framed}

\item If you get an engineering degree, you'll be offered a job as soon as you graduate. \\ 
\begin{framed}
You get an engineering degree and you are not offered a job as soon as you graduate.
\end{framed}

\item If you pass this course, you will graduate this semester. \\ 
\begin{framed}
You will pass this course, and you will not graduate this semester.
\end{framed}

\item If your score is between 12 and 17, you will place into the first course. \\ 
\begin{framed}
Your score is between 12 and 17 and you do not place into the first course.
\end{framed}

\item If your GPA is over 3.7 and you live on campus, you are eligible for this scholarship. \\ \begin{framed}
Your GPA is over 3.7 and you live on campus, but you are not eligible for this scholarship.
\end{framed}
\end{enumerate}

\emph{In problems 9--12, use the distributive laws to write a statement that is logically equivalent to each given statement.}
\begin{enumerate}
\setcounter{enumi}{8}

\item Either the bridge will hold, or those cables will snap and the roadway will crack. \\ 
\begin{framed}
The bridge will hold or those cables will snap, and the bridge will hold or the roadway will crack.
\end{framed}

\item You either meet the job requirements or you don't, but you will not get the job.\\ 
\begin{framed}
Either you meet the job requirements and you will not get the job,\\ or you don't meet them and you will not get the job.
\end{framed}

\item Either get your grades up and get a job, or you won't get a car. \\ 
\begin{framed}
Get your grades up or you won't get a car, and get a job or you won't get a car.
\end{framed}

\item This band is from Texas, and they have either three or four members. \\ 
\begin{framed}
This band is from Texas and they have three members, or they are from Texas and have four members.
\end{framed}
\end{enumerate}

\emph{In problems 13--16, use De Morgan's Laws to write a statement that is logically equivalent to each given statement.}
\begin{enumerate}
\setcounter{enumi}{12}

\item It is not true that North Dakota and East Dakota are both states. \\ 
\begin{framed}
North Dakota is not a state, or East Dakota is not a state.
\end{framed}

\item It is not true that this chapter covers logic and finance. \\ 
\begin{framed}
This chapter does not cover logic or it does not cover finance.
\end{framed}

\item It is not true that this book is entertaining or educational. \\ 
\begin{framed}
This book is not entertaining, and it is not educational.
\end{framed}

\item It is not true that today is Wednesday or Thursday. \\ 
\begin{framed}
Today is not Wednesday, and today is not Thursday.
\end{framed}
\end{enumerate}

\emph{In problems 17--20, use De Morgan's Laws to write the negation of each given statement.}
\begin{enumerate}
\setcounter{enumi}{16}

\item I pay taxes and I vote. \\ 
\begin{framed}
I don't pay taxes or I don't vote.
\end{framed}

\item Either the Packers or the Broncos will win the Super Bowl. \\ 
\begin{framed}
The Packers will not win the Super Bowl, and neither will the Broncos.
\end{framed}

\item Either that smoothie contains green vegetables, or it isn't as healthy as it looks. \\ 
\begin{framed}
That smoothie doesn't contain green vegetables, and it is as healthy as it looks.
\end{framed}

\item Class isn't over, and that clock is fast. \\ 
\begin{framed}
Class is over, or that clock isn't fast.
\end{framed}
\end{enumerate}
\pagebreak

\emph{In problems 21--24, write the contrapositive of each conditional statement.}
\begin{enumerate}
\setcounter{enumi}{20}

\item If he is guilty, he won't testify at his trial. \\ 
\begin{framed}
If he testifies at his trial, he isn't guilty.
\end{framed}

\item If the cat is running, he either spotted a mouse or he spotted a squirrel. \\ 
\begin{framed}
If the cat didn't spot a mouse and he didn't spot a squirrel, he isn't running.
\end{framed}

\item If you do not report for jury duty, or you falsify your information, you will be prosecuted. \\ 
\begin{framed}
If you are not prosecuted, you reported for jury duty and you didn't falsify your information.
\end{framed}

\item If you give your plants water and sunlight, they will survive. \\ 
\begin{framed}
If your plants don't survive, you didn't give them water or you didn't give them sunlight.
\end{framed}
\end{enumerate}