\begin{exercises}
\emph{In problems 1--8, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ You studied.}\\
&q: \textrm{ You passed this course.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\ptwo{If you passed this course, you studied.}
\ptwo{You passed this course if and only if you studied.}

\ptwo{You didn't study if and only if you passed this course.}
\ptwo{You didn't study if you didn't pass this course.}\\ \\

\emph{Write each of the following statements in words.}

\pfour{$p \to q$}
\pfour{$\sim p \leftrightarrow\ \sim q$}
\pfour{$\sim q \to p$}
\pfour{$\sim p \to q$}\\ \\

\emph{In problems 9--16, let $p$, $q$, and $r$ represent the following statements:}
\begin{align*}
&p: \textrm{ The ice cream truck is here.}\\
&q: \textrm{ The pool is open.}\\
&r: \textrm{ It is summertime.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\ptwo{If the ice cream truck is here, and the pool is open, then it is summertime.}
\ptwo{It isn't summertime if the ice cream truck isn't here and the pool isn't open.}

\ptwo{It's summertime if and only if the pool is open.}
\ptwo{If it isn't summertime, the pool isn't open or the ice cream truck isn't here.}\\ \\

\emph{Write each of the following statements in words.}

\pfour{$(p \vee q) \to r$}
\pfour{$r \to (p \wedge q)$}
\pfour{$p \leftrightarrow q$}
\pfour{$p \leftrightarrow (q \wedge r)$}\\ \\

\emph{In problems 17--24, fill in the blanks in each truth table.}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c|}
\hline
$p$ & $q$ & $q \to p$ & $\sim (q \to p)$\\
\hline
& & &\\
T & T & & \\
T & F & & \\
F & T & & \\
F & F & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $\sim q$ & $\sim p \leftrightarrow \sim q$\\
\hline
& & & &\\
T & T & & & \\
T & F & & & \\
F & T & & & \\
F & F & & & \\
\hline
\end{tabular}
\end{center}}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c|}
\hline
$p$ & $q$ & $p \vee q$ & $q \to (p \vee q)$\\
\hline
& & &\\
T & T & & \\
T & F & & \\
F & T & & \\
F & F & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{\small
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $p \wedge q$ & $\sim p$ & $\sim q$ & $\sim p \wedge \sim q$ & $(p \wedge q) \to (\sim p \wedge \sim q)$\\
\hline
& & & & & &\\
T & T & & & & & \\
T & F & & & & & \\
F & T & & & & & \\
F & F & & & & & \\
\hline
\end{tabular}
\end{center}}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $q \to p$ & $p \to q$ & $(q \to p) \wedge (p \to q)$\\
\hline
& & & &\\
T & T & & & \\
T & F & & & \\
F & T & & & \\
F & F & & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $p \to q$ & $\sim q$ & $(p \to q) \wedge \sim q$\\
\hline
& & & &\\
T & T & & & \\
T & F & & & \\
F & T & & & \\
F & F & & & \\
\hline
\end{tabular}
\end{center}}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c | c|}
\hline
$p$ & $q$ & $r$ & $\sim p$ & $\sim p \vee r$ & $(\sim p \vee r) \to q$\\
\hline
& & & & &\\
T & T & T & & & \\
T & F & T & & & \\
F & T & T & & & \\
F & F & T & & & \\
T & T & F & & & \\
T & F & F & & & \\
F & T & F & & & \\
F & F & F & & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $r$ & $\sim p$ & $\sim r$ & $q \to \sim p$ & $\sim r \wedge (q \to \sim p)$\\
\hline
& & & & & &\\
T & T & T & & & & \\
T & F & T & & & & \\
F & T & T & & & & \\
F & F & T & & & & \\
T & T & F & & & & \\
T & F & F & & & & \\
F & T & F & & & & \\
F & F & F & & & & \\
\hline
\end{tabular}
\end{center}}\\ \\

\emph{In problems 25--28, determine if each statement is a tautology or not.}

\ptwo{$p \wedge \sim p$}
\ptwo{$p \vee \sim p$}

\ptwo{$(p \wedge q) \vee (\sim p \wedge q) \leftrightarrow q$}
\ptwo{$(q \to p) \vee (\sim q \to \sim p)$}\\ \\

\emph{In problems 29--30, select the statement that is equivalent to the one given.}

\ptwo{``Either the Cardinals or the Yankees will win the World Series.''
\begin{enumerate}[(a)]
\item If the Cardinals don't win the World Series, the Yankees will win it.
\item The Cardinals and the Yankees will win the World Series.
\item If the Cardinals win the World Series, the Yankees will not win it.
\item If the Yankees win the World Series, the Cardinals will not.
\end{enumerate}}
\ptwo{``If the light is on, someone is home.''
\begin{enumerate}[(a)]
\item The light is on and someone is home.
\item If someone is home, the light is on.
\item Either someone is not home, or the light is on.
\item If someone is not home, the light is not on.
\end{enumerate}}\\ \\

\emph{In problems 31--32, write the converse, inverse, and contrapositive of the given statement.}

\ptwo{If you drive through the field, you see fireflies.}
\ptwo{If you go to office hours, you pass the test.}
\end{exercises}