\begin{exercises}
\emph{In problems 1--8, write the negation of each conditional statement.}

\ptwo{If a fruit is blue, then it is not a banana.}
\ptwo{If the storm comes through, that awning will blow away.}

\ptwo{You'll catch a cold if you don't take Vitamin C.}
\ptwo{You'll get a three percent return on your investment if you invest with us.}

\ptwo{If you get an engineering degree, you'll be offered a job as soon as you graduate.}
\ptwo{If you pass this course, you will graduate this semester.}

\ptwo{If your score is between 12 and 17, you will place into the first course.}
\ptwo{If your GPA is over 3.7 and you live on campus, you are eligible for this scholarship.}\\

\emph{In problems 9--12, use the distributive laws to write a statement that is logically equivalent to each given statement.}

\ptwo{Either the bridge will hold, or those cables will snap and the roadway will crack.}
\ptwo{You either meet the job requirements or you don't, but you will not get the job.}

\ptwo{Either get your grades up and get a job, or you won't get a car.}
\ptwo{This band is from Texas, and they have either three or four members.}\\ \\

\emph{In problems 13--16, use De Morgan's Laws to write a statement that is logically equivalent to each given statement.}

\ptwo{It is not true that North Dakota and East Dakota are both states.}
\ptwo{It is not true that this chapter covers logic and finance.}

\ptwo{It is not true that this book is entertaining or educational.}
\ptwo{It is not true that today is Wednesday or Thursday.}\\ \\

\emph{In problems 17--20, use De Morgan's Laws to write the negation of each given statement.}

\ptwo{I pay taxes and I vote.}
\ptwo{Either the Packers or the Broncos will win the Super Bowl.}

\ptwo{Either that smoothie contains green vegetables, or it isn't as healthy as it looks.}
\ptwo{Class isn't over, and that clock is fast.}\\ \\

\emph{In problems 21--24, write the contrapositive of each conditional statement.}

\ptwo{If he is guilty, he won't testify at his trial.}
\ptwo{If the cat is running, he either spotted a mouse or he spotted a squirrel.}

\ptwo{If you do not report for jury duty, or you falsify your information, you will be prosecuted.}
\ptwo{If you give your plants water and sunlight, they will survive.}
\end{exercises}