\begin{exercises}
\emph{In problems 1--10, determine whether or not each sentence is a statement.}

\ptwo{1024 is the smallest 4-digit number that is a perfect square.}
\ptwo{Doc Brown's time machine requires 1.21 jigowatts of power to operate.}

\ptwo{All elephants are purple.}
\ptwo{Don't do drugs.}

\ptwo{$64=2^5$}
\ptwo{Lizard people run the government.}

\ptwo{What are you holding?}
\ptwo{Don't believe everything you read.}

\ptwo{More soldiers died from sickness than combat in the Revolutionary War.}
\ptwo{This statement is a lie.}\\ \\

\emph{In problems 11--13, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ It will rain this weekend.}\\
&q: \textrm{ We can go to the mall on Saturday.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\pthree{``It won't rain this weekend.''}
\pthree{``Either it will rain this weekend, or we can go to the mall on Saturday.''}
\pthree{``We can't go to the mall on Saturday, but at least it won't rain this weekend.''}\\ \\

\emph{In problems 14--16, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ Interest rates are low.}\\
&q: \textrm{ It is not time to buy a house.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\pthree{``It is time to buy a house; interest rates are low.''}
\pthree{``Interest rates are not low, and it is not time to buy a house.''}
\pthree{``Either it is time to buy a house, or interest rates are not low.''}\\ \\

\emph{In problems 17--22, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ You study hard.}\\
&q: \textrm{ You graduate with honors.}
\end{align*}
\emph{Write each of the following statements in symbolic form.}

\pthree{``You study hard and you graduate with honors.''}
\pthree{``You don't study hard, but you graduate with honors anyway.''}
\pthree{``You study hard, but you still don't graduate with honors.''}

\pthree{``You study hard or you don't graduate with honors.''}
\pthree{``You don't study hard or you graduate with honors anyway.''}
\pthree{``You don't study hard, or you don't graduate with honors.''}\\ \\

\emph{In problems 23--26, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ Kevin Durant will win the MVP award.}\\
&q: \textrm{ The Lakers will win the NBA championship.}
\end{align*}
\emph{Write each of the following statements in words.}

\pfour{$p \vee q$}
\pfour{$p \wedge q$}
\pfour{$\sim p \vee q$}
\pfour{$p\ \wedge \sim q$}\\ \\

\emph{In problems 27--30, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ This is a dog.}\\
&q: \textrm{ This is a mammal.}
\end{align*}
\emph{Write each of the following statements in words.}

\pfour{$\sim p \wedge q$}
\pfour{$p\ \wedge \sim q$}
\pfour{$\sim p\ \vee \sim q$}
\pfour{$\sim q$}\\ \\

\emph{In problems 31--34, let $p$ and $q$ represent the following statements:}
\begin{align*}
&p: \textrm{ Eggs make a good breakfast.}\\
&q: \textrm{ Bacon is not healthy.}
\end{align*}
\emph{Write each of the following statements in words.}

\pfour{$\sim p$}
\pfour{$p\ \vee \sim q$}
\pfour{$\sim p\ \wedge \sim q$}
\pfour{$\sim p \vee q$}\\ \\

\emph{In problems 35--40, evaluate the truth value of the given statement for the given truth values of $p$, $q$, and $r$.}

\ptwo{$p\ \vee \sim q$, where $p$ is T and $q$ is F.}
\ptwo{$\sim p\ \wedge q$, where $p$ is F and $q$ is F.}

\ptwo{$p\ \wedge (q\ \vee \sim r)$, where $p$ is T, $q$ is F, and $r$ is F.}
\ptwo{$\sim (p\ \wedge \sim p)$, where $p$ is F.}

\ptwo{$(p\ \wedge q)\ \vee r$, where $p$ is F, $q$ is T, and $r$ is T.}
\ptwo{$(p\ \vee (\sim p\ \vee q))\ \wedge \sim(q \wedge r)$, where $p$ is F, $q$ is F, and $r$ is T.}\\ \\

\emph{In problems 41--46, fill in the blanks in each truth table.}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $\sim p \wedge q$\\
\hline
& & &\\
T & T & & \\
T & F & & \\
F & T & & \\
F & F & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $p \wedge q$ & $(p \wedge q) \vee \sim p$\\
\hline
& & & &\\
T & T & & & \\
T & F & & & \\
F & T & & & \\
F & F & & & \\
\hline
\end{tabular}
\end{center}}

\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $\sim q$ & $p \vee \sim 	q$ & $\sim p \wedge (p \vee \sim q)$\\
\hline
& & & & & \\
T & T & & & & \\
T & F & & & & \\
F & T & & & & \\
F & F & & & & \\
\hline
\end{tabular}
\end{center}}
\ptwo{
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $\sim p$ & $\sim q$ & $\sim p \vee q$ & $p \wedge \sim q$ & $(p \wedge \sim q) \vee (\sim p \vee q)$\\
\hline
& & & & & & \\
T & T & & & & & \\
T & F & & & & & \\
F & T & & & & & \\
F & F & & & & & \\
\hline
\end{tabular}
\end{center}}

\pone{
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c|}
\hline
$p$ & $q$ & $r$ & $\sim p$ & $\sim r$ & $q\ \vee \sim r$ & $\sim p\ \wedge (q\ \vee \sim r)$\\
\hline
& & & & & & \\
T & T & T & & & & \\
T & F & T & & & & \\
F & T & T & & & & \\
F & F & T & & & & \\
T & T & F & & & & \\
T & F & F & & & & \\
F & T & F & & & & \\
F & F & F & & & & \\
\hline
\end{tabular}
\end{center}}

\pone{{\small
\begin{center}
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c | c|}
\hline
& & & & & & & & & &\\
$p$ & $q$ & $r$ & $\sim p$ & $\sim q$ & $\sim r$ & $\sim p\ \vee q$ & $p\ \vee \sim r$ & $\sim p\ \vee \sim q$ & $(\sim p\ \vee q) \wedge (p\ \vee \sim r)$ & $[(\sim p\ \vee q) \wedge (p\ \vee \sim r)] \wedge (\sim p\ \vee \sim q)$\\
& & & & & & & & & &\\
\hline
& & & & & & & & & &\\
T & T & T & & & & & & & &\\
T & F & T & & & & & & & &\\
F & T & T & & & & & & & &\\
F & F & T & & & & & & & &\\
T & T & F & & & & & & & &\\
T & F & F & & & & & & & &\\
F & T & F & & & & & & & &\\
F & F & F & & & & & & & &\\
\hline
\end{tabular}
\end{center}}}\\ \\

\emph{In problems 47--50, write the negation of each quantified statement.}

\ptwo{All mammals are land-based.}
\ptwo{Some horses are not domesticated.}

\ptwo{Some trains are late.}
\ptwo{All basketball players are not tall.}
\end{exercises}