\begin{enumerate}
\item A jar contains 6 red marbles numbered 1 to 6 and 8 blue marbles numbered 1 to 8. A marble is drawn at random from the jar. Find the probability the marble is red or odd-numbered. \answer{$\dfrac{10}{14}$}
\[\dfrac{6}{14} + \dfrac{7}{14} - \dfrac{3}{14} = \dfrac{10}{14}\]

\item A jar contains 4 red marbles numbered 1 to 4 and 10 blue marbles numbered 1 to 10. A marble is drawn at random from the jar. Find the probability the marble is blue or even-numbered. \answer{$\dfrac{12}{14}$}
\[\dfrac{10}{14} + \dfrac{7}{14} - \dfrac{5}{14} = \dfrac{12}{14}\]

\item You draw one card from a standard 52-card deck. 
\begin{enumerate}
\item What is the probability of selecting a King or a Queen? \answersub{$\dfrac{8}{52}$}
\[P(K\ or\ Q) = P(K) + P(Q) = \dfrac{4}{52} + \dfrac{4}{52} = \dfrac{8}{52}\]
\item What is the probability of selecting a face card or a 10? \answersub{$\dfrac{16}{52}$}
\[P(F\ or\ 10) = P(F) + P(10) = \dfrac{12}{52} + \dfrac{4}{52} = \dfrac{16}{52}\]
\item What is the probability of selecting a spade or a heart? \answersub{$\dfrac{26}{52}$}
\[P(S\ or\ H) = P(S) + P(H) = \dfrac{13}{52} + \dfrac{13}{52} = \dfrac{26}{52}\]
\item What is the probability of selecting a red card or a black card? \answersub{$\dfrac{52}{52}$}
\[P(R\ or\ B) = P(R) + P(B) = \dfrac{26}{52} + \dfrac{26}{52} = \dfrac{52}{52}\]
\end{enumerate}

\item You are dealt a single card from a standard 52-card deck.
\begin{enumerate}
\item Find the probability that you are not dealt a diamond. \answersub{$\dfrac{39}{52}$}
\[P(not\ D) = 1 - P(D) = 1 - \dfrac{13}{52} = \dfrac{39}{52}\]
\item Find the probability that you are not dealt a face card. \answersub{$\dfrac{40}{52}$}
\[P(not\ F) = 1 - P(F) = 1 - \dfrac{12}{52} = \dfrac{40}{52}\]
\item Find the probability that you are not dealt an Ace. \answersub{$\dfrac{48}{52}$}
\[P(not\ A) = 1 - P(A) = 1 - \dfrac{4}{52} = \dfrac{48}{52}\]
\item Find the probability that you are not dealt a jack or a king. \answersub{$\dfrac{44}{52}$}
\[P(not\ (J\ or\ K)) = 1 - (P(J) + P(K)) = 1 - \left(\dfrac{4}{52} + \dfrac{4}{52}\right) = \dfrac{44}{52}\]
\end{enumerate}

\item Consider the following information about a group of 130 FCC students:
\begin{center}
\begin{tabular}{c | c c | c}
\textbf{Gender} & \textbf{Right-handed} & \textbf{Left-handed} & \textbf{Total} \\ \hline 
\textbf{Female} & 58 & 13 & 71\\
\textbf{Male} & 47 & 12 & 59  \\ \hline
\textbf{Total} & 105 & 25 & 130 
\end{tabular}
\end{center}
If one person is randomly selected from the group, what is the probability this student is female or left-handed? \answer{$\dfrac{83}{130}$}
\[P(F\ or\ L) = P(F) + P(L) - P(F\ and\ L) = \dfrac{71}{130} + \dfrac{25}{130} - \dfrac{13}{130} = \dfrac{83}{130}\]

\item The table below shows the number of survey subjects who have received and not received a speeding ticket in the last year, and the color of their car. Find the probability that a randomly chosen person has a red car \emph{or} got a speeding ticket. 
\begin{center}
\begin{tabular}{c | c c | c}
& \textbf{Speeding Ticket} & \textbf{No Speeding Ticket }& \textbf{Total} \\ \hline 
\textbf{Red Car} & 15 & 135 & 150\\
\textbf{Not Red Car} & 45 & 470 & 515  \\ \hline
\textbf{Total} & 60 & 605 & 665
\end{tabular}
\end{center} \text{} \answer{$\dfrac{195}{665}$}
\[P(R\ or\ S) = P(R) + P(S) - P(R\ and\ S) = \dfrac{150}{665} + \dfrac{60}{665} - \dfrac{15}{665} = \dfrac{195}{665}\]

\item Suppose you roll a blue six sided die and a red six sided die, and add their totals.
Find the probability of rolling:
\begin{enumerate}
\item A 7 or 11. \answersub{$\dfrac{8}{36}$}
\[P(7\ or\ 11) = P(7) + P(11) - P(7\ and\ 11) = \dfrac{6}{36} + \dfrac{2}{36} - \dfrac{0}{36} = \dfrac{8}{36}\]
\item An even number or a number less than 6. \answersub{$\dfrac{24}{36}$}
\begin{align*}
P(E\ or\ <6) &= P(E) + P(<6) - P(E\ and\ <6) = 1 - (P(7) + P(9) + P(11))\\
&= 1 - \left(\dfrac{6}{36} + \dfrac{4}{36} + \dfrac{2}{36}\right) = \dfrac{24}{36}
\end{align*}
\item A prime number or a number greater than 5. \answersub{$\dfrac{3}{36}$}
\[P(P\ or\ >5) = P(4) = \dfrac{3}{36}\]
\end{enumerate}

\item A bag contains 4 white counters, 6 black counters, and 1 green counter. What is the
probability of drawing:
\begin{enumerate}
\item A white counter or a green counter? \answersub{$\dfrac{5}{11}$}
\[P(W\ or\ G) = P(W) + P(G) = \dfrac{4}{11} + \dfrac{1}{11} = \dfrac{5}{11}\]
\item A black counter or a green counter? \answersub{$\dfrac{7}{11}$}
\[P(B\ or\ G) = P(B) + P(G) = \dfrac{6}{11} + \dfrac{1}{11} = \dfrac{7}{11}\]
\item Not a green counter? \answersub{$\dfrac{10}{11}$}
\[P(not\ G) = 1 - P(G) = 1 - \dfrac{1}{11} = \dfrac{10}{11}\]
\end{enumerate}

\item  A poll was taken of 14,056 working adults aged 40-70 to determine their level of
education. The participants were classified by sex and by level of education. The results
were as follows.
\begin{center}
\begin{tabular}{c|cc|c}
\textbf{Education Level}     &  \textbf{Male} &  \textbf{Female} &  \textbf{Total} \\ \hline
\textbf{High School or Less} & 3141  &2434     &  5575  \\ 
\textbf{Bachelor's Degree}   & 3619  &3761     &  7380  \\
\textbf{Master's Degree}     & 534   &472      &  1006  \\
\textbf{Ph.D.}               & 52    &43       &  95    \\ \hline
\textbf{Total} 							& 7346  &6710     & 14,056 \\ 
\end{tabular}
\end{center}
A person is selected at random. Compute the following probabilities:
\begin{enumerate}
\item The probability that the selected person does not have a Ph.D. \answersub{$\dfrac{13,961}{14,056}$}
\[P(not\ P) = 1 - P(P) = 1 - \dfrac{95}{14,056} = \dfrac{13,961}{14,056}\]
\item The probability that the selected person does not have a Master's degree. \answersub{$\dfrac{13,961}{14,056}$}
\[P(not\ M) = 1 - P(M) = 1 - \dfrac{1006}{14,056} = \dfrac{13,050}{14,056}\]
\item The probability that the selected person is female or has a Master's degree. \answersub{$\dfrac{7244}{14,056}$}
\[P(F\ or\ M) = P(F) + P(M) - P(F\ and\ M) = \dfrac{6710}{14,056} + \dfrac{1006}{14,056} - \dfrac{472}{14,056} = \dfrac{7244}{14,056}\]
\item The probability that the selected person is male or has a Ph.D. \answersub{$\dfrac{7389}{14,056}$}
\[P(M\ or\ P) = P(M) + P(P) - P(M\ and\ P) = \dfrac{7346}{14,056} + \dfrac{95}{14,056} - \dfrac{52}{14,056} = \dfrac{7389}{14,056}\]
\end{enumerate}
\end{enumerate}