\begin{enumerate}
\item A fair die is rolled. Find the probability of getting 4. \answer{$\dfrac{4}{6}$}

\item A fair die is rolled. Find the probability of getting less than 3. \answer{$\dfrac{2}{6}$}

\item A fair die is rolled. Find the probability of getting at least 5. \answer{$\dfrac{2}{6}$}

\item You have a bag with 20 cherries, 14 sweet and 6 sour. If you pick a cherry at random, what is the probability that it will be sweet? \answer{$\dfrac{14}{20}$}

\item A ball is drawn randomly from a jar that contains 6 red balls, 2 white balls, and 5 yellow balls. Find the probability of drawing a white ball. \answer{$\dfrac{2}{13}$}

\item Suppose you write each letter of the alphabet on a different slip of paper and put the slips into a hat. What is the probability of drawing one slip of paper from the hat at random and getting a consonant?  Note that a consonant is any letter that is not a vowel (\emph{a, e, i, o,} or \emph{u}). \answer{$\dfrac{21}{26}$}

\item In a survey, 205 people indicated they prefer cats, 160 indicated they prefer dogs, and 40 indicated they don't enjoy either pet. Find the probability that if a person is chosen at random, they prefer cats. \answer{$\dfrac{205}{405}$}

\item A group of people were asked if they had run a red light in the last year. 150 responded ``yes'' and 185 responded ``no.'' Find the probability that if a person is chosen at random, they have run a red light in the last year. \answer{$\dfrac{150}{335}$}

\item A U.S. roulette wheel has 38 pockets: 1 through 36, 0, and 00. 18 are black, 18 are red, and 2 are green. A play has a dealer spin the wheel and a small ball in opposite directions. As the ball slows to stop, it can land with equal probability on the 38 slots. Find the probability of the ball landing on green. \answer{$\dfrac{2}{38}$}

\item A glass jar contains 6 red, 5 green, 8 blue and 3 yellow marbles. If a single marble is
chosen at random from the jar, what is the probability of choosing
\begin{enumerate}[(a)]
\item a red marble? \answersub{$\dfrac{6}{22}$}
\item a green marble? \answersub{$\dfrac{5}{22}$}
\item a blue marble? \answersub{$\dfrac{8}{22}$}
\end{enumerate}

\item Lisa has a large bag of coins. After counting the coins, she recorded the counts in the table below. She then decided to draw some coins at random, replacing each coin before the next draw.
\begin{center}
\begin{tabular}{c c c c}
\textbf{Quarters} & \textbf{Nickels} & \textbf{Dimes} & \textbf{Pennies} \\
\hline 
27 & 18 & 34 & 21
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item What is the probability that Lisa obtains a quarter on the first draw? \answersub{$\dfrac{27}{100}$}
\item What is the probability that Lisa obtains a penny or a dime on the second draw? \answersub{$\dfrac{55}{100}$}
\item What is the probability that Lisa obtains at most 10 cents worth of money on the third draw? \answersub{$\dfrac{73}{100}$}
\item What is the probability that Lisa does not get a nickel on the fourth draw? \answersub{$\dfrac{82}{100}$}
\item What is the probability that Lisa obtains at least 10 cents worth of money on the fifth draw? \answersub{$\dfrac{61}{100}$}
\end{enumerate}
		
\item Suppose you roll a pair of six-sided dice. 
\begin{enumerate}[(a)]
\item List all possible outcomes of this experiment. \answersub{listed below (36 total possibilities}
\begin{align*}
S = \{&(1,1), (1,2), (1,3), (1,4), (1,5), (1,6)\\
&(2,1), (2,2), (2,3), (2,4), (2,5), (2,6)\\
&(3,1), (3,2), (3,3), (3,4), (3,5), (3,6)\\
&(4,1), (4,2), (4,3), (4,4), (4,5), (4,6)\\
&(5,1), (5,2), (5,3), (5,4), (5,5), (5,6)\\
&(6,1), (6,2), (6,3), (6,4), (6,5), (6,6)\}
\end{align*}
\item What is the probability that the sum of the numbers on your dice is exactly 6? \answersub{$\dfrac{5}{36}$}
\item What is the probability that the sum of the numbers on your dice is at most 4? \answersub{$\dfrac{6}{36}$}
\item What is the probability that the sum of the numbers on your dice is at least 9? \answersub{$\dfrac{10}{36}$}
\end{enumerate}
\pagebreak

\item I asked my Facebook friends to complete a two-question survey. They answered the following questions:
Which beverage do you prefer in the morning: coffee or tea?  What is your gender? I summarized the results in following table:
\begin{center}
\begin{tabular}{c | c c | c}
& \textbf{Coffee} & \textbf{Tea} & \textbf{Total} \\
\hline
\textbf{Female} & 37 & 24 & \textbf{61} \\
\textbf{Male} & 22 & 31 & \textbf{53} \\
\hline
\textbf{Total } & \textbf{59 }& \textbf{55} & \textbf{114}
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item What is the probability that I select a friend who prefers coffee? \answersub{$\dfrac{59}{114}$}
\item What is the probability that I select a friend who is female? \answersub{$\dfrac{61}{114}$}
\item What is the probability that I select a friend who is male and prefers tea? \answersub{$\dfrac{31}{114}$}
\end{enumerate}	

\item A poll was taken of 14,056 working adults aged 40-70 to determine their level of
education. The participants were classified by sex and by level of education. The results
were as follows.
\begin{center}
\begin{tabular}{c|cc|c}
\textbf{Education Level}     &  \textbf{Male} &  \textbf{Female} &  \textbf{Total} \\ \hline
\textbf{High School or Less} & 3141  &2434     &  5575  \\ 
\textbf{Bachelor's Degree}   & 3619  &3761     &  7380  \\
\textbf{Master's Degree}     & 534   &472      &  1006  \\
\textbf{Ph.D.}               & 52    &43       &  95    \\ \hline
\textbf{Total} 							& 7346  &6710     & 14,056 \\ 
\end{tabular}
\end{center}
A person is selected at random. Compute the following probabilities:
\begin{enumerate}[(a)]
\item The probability that the selected person is a male \answersub{$\dfrac{7346}{14,056}$}
\item The probability that the selected person does not have a Ph.D. \answersub{$\dfrac{13,961}{14,056}$}
\item The probability that the selected person has a Master's degree \answersub{$\dfrac{1006}{14,056}$}
\item The probability that the selected person is female and has a Master's degree \answersub{$\dfrac{472}{14,056}$}
\end{enumerate}	
\end{enumerate}