\emph{In problems 1--4, use your calculator to evaluate the given expression.}
\begin{enumerate}
\item $12!$ \answer{479,001,600}

\item $\dfrac{5!}{3!}$ \answer{20}

\item $_8P_3$ \answer{336}

\item $_{22}C_4$ \answer{7315}

\item A license plate is to have the following form: three letters followed by three
numbers. An example of a license plate would be MTH 314. How many different license plates can be made, assuming that letters and numbers can be reused? \answer{17,576,000}
\[26 \cdot 26 \cdot 26 \cdot 10 \cdot 10 \cdot 10 = 17,576,000\]

\item A bride is choosing a dress for her bridesmaids. If there are 4 styles, 3 colors, and 6 fabrics, in how many ways can she select the bridesmaids' dress? \answer{72}
\[4 \cdot 3 \cdot 6 = 72\]

\item A quiz consists of 5 true-or-false questions. In how many ways can a student answer the quiz? \answer{32}
\[2^5 = 32\]

\item A boy owns 2 pairs of pants, 3 shirts, 8 ties, and 2 jackets. How many different outfits can he wear to school if he must wear one of each item? \answer{96}
\[2 \cdot 3 \cdot 8 \cdot 2 = 96\]

\item Lisa is shopping for a new car. She has the following decisions to make: type (sedan, SUV, pick-up truck), make (domestic or import), color (black, white, silver). In how many ways can she select her new vehicle? \answer{18}
\[3 \cdot 2 \cdot 3 = 18\]

\item A social security number contains nine digits, such as 999-04-6756. How many different social security numbers can be formed? Do you think we will ever run out? \answer{1,000,000,000}
\[10^9 = 1,000,000,000\]

%%% Permutations %%%
\item How many different ways can the letters of the word MATH be rearranged to form a four-letter code word? \answer{24}
\[4 \cdot 3 \cdot 2 \cdot 1 = 24\]

\item How many ways can we select five door prizes from seven different ones and distribute them among five people? \answer{2520}
\[7\ nPr\ 5 = 2520\]

\item Eight sprinters have made it to the Olympic finals in the 100-meter race. In how many different ways can the gold, silver, and bronze medals be awarded? \answer{336}
\[8\ nPr\ 3 = 336\]

\item At a charity benefit with 25 people in attendance, three \$50 gift certificates are given away as door prizes. Assuming no person receives more than one prize, how many different ways can the gift certificates be awarded? \answer{2300}
\[25\ nCr\ 3 = 2300\]

\item Ten bands are to perform at a weekend festival. How many different ways
are there to schedule their appearances? \answer{3,628,800}
\[10! = 3,628,800\]

\item There are seven books in the Harry Potter series. In how many ways can you arrange the books on your shelf? \answer{5040}
\[7! = 5040\]

%%% Combinations %%%
\item How many different four-card hands can be dealt from a 52 card deck? \answer{270,725}
\[52\ nCr\ 4 = 270,725\]

\item There are 40 runners in a race, and no ties. In how many ways can the first three finishers be chosen from the 40 runners, regardless of how they are arranged? \answer{9880}
\[40\ nCr\ 3 = 9880\]

\item There are 24 students in a MATH 120 class at FCC. How many ways are there to select 5 students for a group project? \answer{42,504}
\[24\ nCr\ 5 = 42,504\]

\item A local children's center has 55 kids, and 6 are selected to take a picture for the center's advertisement. How many ways are there to select 6 children for the picture? \answer{28,989,675}
\[55\ nCr\ 6 = 28,989,675\]
\pagebreak

%%% Probability problems %%%
\item In a lottery game, a player picks six numbers from 1 to 48. If exactly 4 of the 6 numbers match those drawn, the player wins third prize. What is the probability of winning this prize? \answer{$\dfrac{12,915}{12,271,512} = 0.0011$}
\begin{center}
Total possibilities: choose 6 numbers from 48: $48\ nCr\ 6 = 12,271,512$\\
Ways to win: choose 4 of the 6 to match, 2 of the 42 to not match: $(6\ nCr\ 4) \cdot (42\ nCr\ 2) = 12,915$\\
Answer: $\dfrac{12,915}{12,271,512} = 0.0011$
\end{center}

\item Compute the probability of randomly drawing five cards from a deck and getting 3 Aces and 2 Kings. \answer{$\dfrac{24}{2,598,960} = 0.000009$}
\begin{center}
Total possibilities: choose 5 cards from 52: $52\ nCr\ 5 = 2,598,960$\\
Ways to win: choose 3 out of 4 Aces, 2 out of 4 Kings: $(4\ nCr\ 3) \cdot (4\ nCr\ 2) = 24$\\
Answer: $\dfrac{24}{2,598,960} = 0.000009$
\end{center}

\item You own 16 CDs. You want to randomly arrange 5 of them in a CD rack. What is the probability that the rack ends up in alphabetical order? \answer{$\dfrac{1}{120} = 0.0083$}
\begin{center}
\textbf{NOTE: the 16 is irrelevant.  Once you select 5 CDs, there is one correct alphabetical ordering out of all possible ways to arrange the 5.}\\
Total possibilities: arrange 5: $5! = 120$\\
Ways to arrange alphabetically: 1\\
Answer: $\dfrac{1}{120} = 0.0083$
\end{center}

\item A jury pool consists of 27 people, 14 men and 13 women. Compute the probability that a randomly selected jury of 12 people is all male. \answer{$\dfrac{91}{17,383,860} = 0.000005$}
\begin{center}
Total possibilities: select 12 people from 27: $27\ nCr\ 12 = 17,383,860$\\
All male: select 12 people out of 14: $14\ nCr\ 12 = 91$\\
Answer: $\dfrac{91}{17,383,860} = 0.000005$
\end{center}

\item Compute the probability that a 5-card poker hand is dealt to you that contains all hearts. \answer{$\dfrac{1287}{2,598,960} = 0.0005$}
\begin{center}
Total possibilities: select 5 cards from 52: $52\ nCr\ 5 = 2,598,960$\\
All hearts: select 5 cards out of 13: $13\ nCr\ 5 = 1287$\\
Answer: $\dfrac{1287}{2,598,960} = 0.0005$
\end{center}

\item Compute the probability that a 5-card poker hand is dealt to you that contains four Aces. \answer{$\dfrac{48}{2,598,960} = 0.00002$}
\begin{center}
Total possibilities: select 5 cards from 52: $52\ nCr\ 5 = 2,598,960$\\
Four Aces: select 4 cards out of 4 and 1 card out of 48: $(4\ nCr\ 4) \cdot (48\ nCr\ 1) = 48$\\
Answer: $\dfrac{48}{2,598,960} = 0.00002$
\end{center}

\item A jury pool has 18 men and 21 women, from which 12 jurors will be selected. Assuming that
each person is equally likely to be selected and that the jury is selected at random, find the
probability the jury consists of
\begin{enumerate}[(a)]
\item all men \answersub{$\dfrac{18,564}{3,910,797,436} = 0.000005$}
\[\dfrac{18\ nCr\ 12}{39\ nCr\ 12} = \dfrac{18,564}{3,910,797,436} = 0.000005\]
\item all women \answersub{$\dfrac{293,930}{3,910,797,436} = 0.00008$}
\[\dfrac{21\ nCr\ 12}{39\ nCr\ 12} = \dfrac{293,930}{3,910,797,436} = 0.00008\]
\item 8 men and 4 women \answersub{$\dfrac{261,891,630}{3,910,797,436} = 0.0670$}
\[\dfrac{(18\ nCr\ 8) \cdot (21\ nCr\ 4)}{39\ nCr\ 12} = \dfrac{261,891,630}{3,910,797,436} = 0.0670\]
\item 6 men and 6 women \answersub{$\dfrac{1,007,356,896}{3,910,797,436} = 0.2576$}
\[\dfrac{(18\ nCr\ 6) \cdot (21\ nCr\ 6)}{39\ nCr\ 12} = \dfrac{1,007,356,896}{3,910,797,436} = 0.2576\]
\end{enumerate}

\item A race consisted of 8 women and 10 men. What is the probability that the top 3
finishers were:
\begin{enumerate}[(a)]
\item all men \answersub{$\dfrac{120}{816} = 0.1471$}
\[\dfrac{10\ nCr\ 3}{18\ nCr\ 3} = \dfrac{120}{816} = 0.1471\]
\item all women \answersub{$\dfrac{56}{816} = 0.0686$}
\[\dfrac{8\ nCr\ 3}{18\ nCr\ 3} = \dfrac{56}{816} = 0.0686\]
\item 2 men and 1 woman \answersub{$\dfrac{360}{816} = 0.4412$}
\[\dfrac{(10\ nCr\ 2) \cdot (8\ nCr\ 1)}{18\ nCr\ 3} = \dfrac{360}{816} = 0.4412\]
\item 1 man and 2 women \answersub{$\dfrac{280}{816} = 0.3431$}
\[\dfrac{(10\ nCr\ 1) \cdot (8\ nCr\ 2)}{18\ nCr\ 3} = \dfrac{280}{816} = 0.3431\]
\end{enumerate}
\end{enumerate}