\begin{enumerate}
\item You have a box of chocolates that contains 50 pieces, of which 30 are solid chocolate,
15 are filled with cashews and 5 are filled with cherries. All the candies look exactly alike.
You select a piece, eat it, select a second piece, eat it, and finally eat one last piece. Find the probability of selecting a solid chocolate followed by two cherry-filled chocolates. \answer{$\dfrac{600}{117,600} = 0.0051$}
\[P(S,C,C) = P(S) \cdot P(C | S) \cdot P(C | S,C) = \dfrac{30}{50} \cdot \dfrac{5}{49} \cdot {4}{48} = \dfrac{600}{117,600} = 0.0051\]

\item You roll a fair six-sided die twice. Find the probability of rolling a 6 the first time
and a number greater than 2 the second time. \answer{$\dfrac{4}{36} = 0.1111$}
\[P(6,>2) = P(6) \cdot P(>2) = \dfrac{1}{6} \cdot \dfrac{4}{6} = \dfrac{4}{36} = 0.1111\]

\item A math class consists of 25 students, 14 female and 11 male. Three students are selected at random, one at a time, to participate in a probability experiment. Compute the probability that:
\begin{enumerate}[(a)]
\item A male is selected, then two females. \answersub{$\dfrac{2002}{13,800} = 0.1451$}
\[P(M,F,F) = P(M) \cdot P(F | M) \cdot P(F | M,F) = \dfrac{11}{25} \cdot \dfrac{14}{24} \cdot \dfrac{13}{23} = \dfrac{2002}{13,800} = 0.1451\]
\item A female is selected, then two males. \answersub{$\dfrac{1540}{13,800} = 0.1116$}
\[P(F,M,M) = P(F) \cdot P(M | F) \cdot P(M | F,M) = \dfrac{14}{25} \cdot \dfrac{11}{24} \cdot \dfrac{10}{23} = \dfrac{1540}{13,800} = 0.1116\]
\item Two females are selected, then one male. \answersub{$\dfrac{2002}{13,800} = 0.1451$}
\[P(F,F,M) = P(F) \cdot P(F | F) \cdot P(M | F,F) = \dfrac{14}{25} \cdot \dfrac{13}{24} \cdot \dfrac{11}{23} = \dfrac{2002}{13,800} = 0.1451\]
\item Three males are selected. \answersub{$\dfrac{990}{13,800} = 0.0717$}
\[P(M,M,M) = P(M) \cdot P(M | M) \cdot P(M | M,M) = \dfrac{11}{25} \cdot \dfrac{10}{24} \cdot \dfrac{9}{23} = \dfrac{990}{13,800} = 0.0717\]
\item Three females are selected. \answersub{$\dfrac{2184}{13,800} = 0.1583$}
\[P(F,F,F) = P(F) \cdot P(F | F) \cdot P(F | F,F) = \dfrac{14}{25} \cdot \dfrac{13}{24} \cdot \dfrac{12}{23} = \dfrac{2184}{13,800} = 0.1583\]
\end{enumerate}
\pagebreak

\item A large cooler contains the following drinks: 6 lemonade, 8 Sprite, 15 Coke, and 7 root beer. You randomly pick two cans, one at a time (without replacement).
\begin{enumerate}[(a)]
\item What is the probability that you get 2 cans of Sprite? \answersub{$\dfrac{56}{1260} = 0.0444$}
\[P(S,S) = P(S) \cdot P(S | S) = \dfrac{8}{36} \cdot \dfrac{7}{35} = \dfrac{56}{1260} = 0.0444\]
\item What is the probability that you do not get 2 cans of Coke? \answersub{$\dfrac{1050}{1260} = 0.8333$}
\[P(not\ C,C) = 1 - P(C) \cdot P(C | C) = 1 - \dfrac{15}{36} \cdot \dfrac{14}{35} = 1 - \dfrac{210}{1260} = \dfrac{1050}{1260} = 0.8333\]
\item What is the probability that you get either 2 root beer or 2 lemonade? \answersub{$\dfrac{72}{1260} = 0.0571$}
\begin{align*}
P(R,R\ or\ L,L) &= P(R,R) + P(L,L) = P(R) \cdot P(R | R) + P(L) \cdot P(L | L)\\
&= \dfrac{7}{36} \cdot \dfrac{6}{35} + \dfrac{6}{36} \cdot \dfrac{5}{35} = \dfrac{72}{1260} = 0.0571
\end{align*}
\item What is the probability that you get one can of Coke and one can of Sprite? \answersub{$\dfrac{240}{1260} = 0.1905$}
\[P(C,S) + P(S,C) = P(C) \cdot P(S | C) + P(S) \cdot P(C | S) = \dfrac{15}{36} \cdot \dfrac{8}{35} + \dfrac{8}{36} \cdot \dfrac{15}{35} = \dfrac{240}{1260} = 0.1905\]
\item What is the probability that you get two drinks of the same type? \answersub{$\dfrac{338}{1260} = 0.2683$}
\[P(L,L) + P(S,S) + P(C,C) + P(R,R) = \dfrac{6}{36} \cdot \dfrac{5}{35} + \dfrac{8}{36} \cdot \dfrac{7}{35} + \dfrac{15}{36} \cdot \dfrac{14}{35} + \dfrac{7}{36} \cdot \dfrac{6}{35} = \dfrac{338}{1260} = 0.2683\]
\end{enumerate}

\item My top drawer contains different colored socks: 14 are white, 10 are black, 6 are pink, and 4 are blue. All socks in the drawer are loose. Every morning I randomly select 2 socks, one at a time. Calculate the following probabilities, giving both fraction and decimal answers, rounding to 4 decimal places:
\begin{enumerate}[(a)]
\item What is the probability that I get a blue pair of socks? \answersub{$\dfrac{12}{1122} = 0.0107$}
\[\dfrac{4}{34} \cdot \dfrac{3}{33} = \dfrac{12}{1122} = 0.0107\]
\item What is the probability that I do not get a blue pair of socks? \answersub{$\dfrac{1110}{1122} = 0.9893$}
\[1 - \dfrac{12}{1122} = \dfrac{1110}{1122} = 0.9893\]
\item What is the probability that I either get a white pair or a blue pair of socks? \answersub{$\dfrac{194}{1122} = 0.1729$}
\[\dfrac{14}{34} \cdot \dfrac{13}{33} + \dfrac{12}{1122} = \dfrac{194}{1122} = 0.1729\]
\item What is the probability that I get one black sock and one white sock? \answersub{$\dfrac{280}{1122} = 0.2496$}
\[\dfrac{10}{34} \cdot \dfrac{14}{33} + \dfrac{14}{34} \cdot \dfrac{10}{33} = \dfrac{280}{1122} = 0.2496\]
\end{enumerate}

\item Suppose a math class contains 25 students, 14 females (three of whom speak French) and 11 males (two of whom speak French). Compute the probability that a randomly selected student speaks French, given that the student is female. \answer{$\dfrac{3}{14} = 0.2143$}

\item Suppose a math class contains 30 students, 18 females (four of whom speak French) and 12 males (three of whom speak French). Compute the probability that a randomly selected student is male, given that the student speaks French. \answer{$\dfrac{3}{7} = 0.4286$}

\item A certain virus infects one in every 400 people. A test used to detect the virus in a person is positive 90\% of the time if the person has the virus and 10\% of the time if the person does not have the virus.% Let A be the event ``the person is infected'' and B be the event ``the person tests positive.''
\begin{enumerate}[(a)]
\item Find the probability that a person has the virus given that they have tested positive. \answersub{$0.0221 = 2.21\%$}
\[P(virus | positive) = \dfrac{P(virus\ AND\ positive)}{P(positive)} = \dfrac{0.9}{0.9 + 39.9} = 0.0221\]
\item Find the probability that a person does not have the virus given that they test negative. \answersub{$0.9997 = 99.97\%$}
\[P(no\ virus | negative) = \dfrac{P(no\ virus\ AND\ negative)}{P(negative)} = \dfrac{399 \cdot 0.9}{399 \cdot 0.9 + 0.1} = 0.9997\] 
\end{enumerate}

\item A poll was taken of 14,056 working adults aged 40-70 to determine their level of
education. The participants were classified by sex and by level of education. The results
were as follows.
\begin{center}
\begin{tabular}{c|cc|c}
\textbf{Education Level}     &  \textbf{Male} &  \textbf{Female} &  \textbf{Total} \\ \hline
\textbf{High School or Less} & 3141  &2434     &  5575  \\ 
\textbf{Bachelor's Degree}   & 3619  &3761     &  7380  \\
\textbf{Master's Degree}     & 534   &472      &  1006  \\
\textbf{Ph.D.}               & 52    &43       &  95    \\ \hline
\textbf{Total} 							& 7346  &6710     & 14,056 \\ 
\end{tabular}
\end{center}
A person is selected at random. Compute the following probabilities:
\begin{enumerate}
\item The probability that the selected person is male, given he has a Master's degree. \answersub{$\dfrac{534}{1006} = 0.5308$}
\item The probability that the selected person does not have a Master's degree, given it is a male. \answersub{$\dfrac{6812}{7346} = 0.9273$}
\item The probability that the selected person is female, given that she has a Bachelor's degree. \answersub{$\dfrac{3761}{7380} = 0.0.5096$}
\item The probability that the selected person has a Ph.D, given it is a female. \answersub{$\dfrac{43}{6710} = 0.0064$}
\end{enumerate}
\end{enumerate}