\begin{exercises}
\ptwo{Marko currently has 20 tulips in his yard.  Each year he plants 5 more.
\begin{enumerate}[(a)]
\item Find a linear model of the form $P_t = P_0 + dt$ to describe the number of tulips Marko has at a given point in time.
\item How many tulips will Marko have in 7 years?
\item When will Marko have 65 tulips?
\end{enumerate}}
\ptwo{Pam is a DJ.  Every week she buys 3 new albums to add to her collection.  She currently owns 450 albums.
\begin{enumerate}[(a)]
\item Find a linear model of the form $P_t = P_0 + dt$ to describe the number of albums Pam has at a given point in time.
\item How many albums will Pam have in 11 weeks?
\item When will Pam have 489 albums?
\end{enumerate}}

\ptwo{A store did \$40,000 in sales in 2016, and \$62,000 in 2018.
\begin{enumerate}[(a)]
\item Assuming the store's sales are growing linearly, find the growth rate $d$.
\item Write a linear model to describe this store's sales from 2016 onward.
\item Predict the store's sales in 2025.
\item When do you expect the store's sales to exceed \$100,000?
\end{enumerate}}
\ptwo{A small town had 340 homes in 2010, and by 2020, this had grown to 375.
\begin{enumerate}[(a)]
\item Assuming the number of homes is growing linearly, find the growth rate $d$.
\item Write a linear model to describe the number of homes in this town from 2010 onward.
\item Predict how many homes there will be in this town in 2034.
\item When do you expect the number of homes to exceed 500?
\end{enumerate}}

\ptwo{A population of beetles is growing according to a linear growth model.  Initially, there were 13 beetles, and 8 weeks later, there were 42 beetles.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of beetles over time, using weeks as the unit of time.
\item How many beetles are there expected to be 14 weeks after the initial point?
\item When do you expect the number of beetles to exceed 70?
\end{enumerate}}
\ptwo{The number of streetlights in a town is growing linearly.  Four months ago there were 130 lights.  Now there are 146 lights.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of streetlights in the town over time, using months as the unit of time.
\item How many streetlights are expected a year from \emph{now} (not a year from the beginning)?
\item When do you expect the number of streetlights to exceed 200?
\end{enumerate}}

\ptwo{In 1990, there were 112 nuclear power plants in the U.S.  By 2019, this number had fallen to 96.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of nuclear power plants from 1990 onward.
\item Using this linear model, predict the number of nuclear power plants in the U.S. in 2030.
\item When do you expect the number of nuclear power plants to reach 80?
\end{enumerate}}
\ptwo{In 1990, approximately 1,820,000 violent crimes were reported in the U.S.  By 2019, this number had fallen to approximately 1,200,000.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of violent crimes in the U.S. from 1990 onward.
\item Using this linear model, predict the number of violent crimes in 2040.
\item When do you expect the number of violent crimes to reach 1,000,000?
\end{enumerate}}
\vfill
\pagebreak

\ptwo{The table below shows the average annual cost of health insurance for a single individual, from 1999 to 2019, according to the Kaiser Family Foundation.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Cost}\\
\hline
 & \\
1999 & \$2,196\\
2000 & \$2,471\\
2001 & \$2,689\\
2002 & \$3,083\\
2003 & \$3,383\\
2004 & \$3,695\\
2005 & \$4,024\\
2006 & \$4,242\\
2007 & \$4,479\\
2008 & \$4,704\\
2009 & \$4,824\\
2010 & \$5,049\\
2011 & \$5,429\\
2012 & \$5,615\\
2013 & \$5,884\\
2014 & \$6,025\\
2015 & \$6,251\\
2016 & \$6,435\\
2017 & \$6,690\\
2018 & \$6,896\\
2019 & \$7,186
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Using only the data from the first and last years, build a linear model to describe the cost of individual health insurance from 1999 onward.
\item Using this linear model, predict the cost of insurance in 2030.
\item According to this model, when do you expect the cost of individual insurance to reach \$12,000?
\item Using a calculator or spreadsheet program, build a linear regression model to describe the cost of individual insurance from 1999 onward.
\item Using the regression model, predict the cost of insurance in 2030.
\item According to the regression model, when do you expect the cost of individual insurance to reach \$12,000?
\end{enumerate}}
\ptwo{The table below shows the average annual cost of health insurance for a family, from 1999 to 2019, according to the Kaiser Family Foundation.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Cost}\\
\hline
 & \\
1999 & \$5,791\\
2000 & \$6,438\\
2001 & \$7,061\\
2002 & \$8,003\\
2003 & \$9,068\\
2004 & \$9,950\\
2005 & \$10,880\\
2006 & \$11,480\\
2007 & \$12,106\\
2008 & \$12,680\\
2009 & \$13,375\\
2010 & \$13,770\\
2011 & \$15,073\\
2012 & \$15,745\\
2013 & \$16,351\\
2014 & \$16,834\\
2015 & \$17,545\\
2016 & \$18,142\\
2017 & \$18,764\\
2018 & \$19,616\\
2019 & \$20,576
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Using only the data from the first and last years, build a linear model to describe the cost of family health insurance from 1999 onward.
\item Using this linear model, predict the cost of insurance in 2025.
\item According to this model, when do you expect the cost of family insurance to reach \$30,000?
\item Using a calculator or spreadsheet program, build a linear regression model to describe the cost of family insurance from 1999 onward.
\item Using the regression model, predict the cost of insurance in 2025.
\item According to the regression model, when do you expect the cost of family insurance to reach \$30,000?
\end{enumerate}}
\end{exercises}