\begin{exercises}
\ptwo{One hundred trout are seeded into a lake.  Absent constraint, their population will grow by 70\% a year.  If the lake can sustain a maximum of 2000 trout, use a logistic growth model to estimate the number of trout after 2 years.}
\ptwo{Ten blackberry plants started growing in a yard.  Absent constraint, blackberries will spread by 200\% a month.  If the yard can only sustain 50 plants, use a logistic growth model to estimate the number of plants after 2 months.}

\ptwo{A certain community consists of 1000 people, and one individual has a particularly contagious strain of influenza.  Assuming the community has not had vaccination shots and are all susceptible, the spread of the disease in the community is modeled by \[P_t = \dfrac{1000}{1+999e^{-0.3t}}\] where $P_t$ is the number of people who have contracted the flu after $t$ days.
\begin{enumerate}[(a)]
\item How many people have contracted the flu after 10 days?  Round your answer to the nearest whole number.
\item What is the carrying capacity for this model?  Does this make sense?
\item How many days will it take for 750 people to contract the flu?  Round your answer to the nearest whole number.
\end{enumerate}}
\ptwo{A herd of 20 white-tailed deer is introduced to a coastal island where there had been no deer before.  Their population is predicted to increase according to \[P_t = \dfrac{100}{1+4e^{-0.14t}}\] where $P_t$ is the number of deer expected in the herd after $t$ years.
\begin{enumerate}[(a)]
\item How many deer will be present after 2 years?  Round your answer to the nearest whole number.
\item What is the carrying capacity for this model?
\item How many years will it take for the herd to grow to 50 deer?  Round your answer to the nearest whole number.
\end{enumerate}}

\ptwo{The table below shows the population of California from 2010 to 2019.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Population (millions)}\\
\hline
 & \\
2010 & 37.3\\
2011 & 37.6\\
2012 & 38.0\\
2013 & 38.3\\
2014 & 38.6\\
2015 & 38.9\\
2016 & 39.2\\
2017 & 39.4\\
2018 & 39.5\\
2019 & 39.5
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Use a graphing calculator to build a logistic regression model, letting $t=0$ in 2010.
\item What does this model predict that the population of California will be in 2025?
\item When does this model predict that California's population will reach 40 million?
\item According to this model, what is the carrying capacity for California's population?
\end{enumerate}}
\ptwo{The table below shows the population of Florida from 2010 to 2019.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Population (millions)}\\
\hline
 & \\
2010 & 18.7\\
2011 & 19.1\\
2012 & 19.3\\
2013 & 19.6\\
2014 & 19.9\\
2015 & 20.2\\
2016 & 20.6\\
2017 & 21.0\\
2018 & 21.2\\
2019 & 21.5
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Use a graphing calculator to build a logistic regression model, letting $t=0$ in 2010.
\item What does this model predict that the population of Florida will be in 2030?
\item When does this model predict that Florida's population will reach 23 million?
\item According to this model, what is the carrying capacity for Florida's population?
\end{enumerate}}
\end{exercises}