\begin{enumerate}
\item Marko currently has 20 tulips in his yard.  Each year he plants 5 more.
\begin{enumerate}[(a)]
\item Find a linear model of the form $P_t = P_0 + dt$ to describe the number of tulips Marko has at a given point in time. \answersub{$P_t = 20 + 5t$}
\item How many tulips will Marko have in 7 years? \answersub{55 tulips}
\[P_t = 20 + 5(7) = 55\]
\item When will Marko have 65 tulips? \answersub{In 9 years}
\[65 = 20 + 5t \longrightarrow 45 = 5t \longrightarrow t = 9\]
\end{enumerate}

\item Pam is a DJ.  Every week she buys 3 new albums to add to her collection.  She currently owns 450 albums.
\begin{enumerate}[(a)]
\item Find a linear model of the form $P_t = P_0 + dt$ to describe the number of albums Pam has at a given point in time. \answersub{$P_t = 450 + 3t$}
\item How many albums will Pam have in 11 weeks? \answersub{483 albums}
\[P_t = 450 + 3(11) = 483\]
\item When will Pam have 489 albums? \answersub{In 13 weeks}
\[489 = 450 + 3t \longrightarrow 39 = 3t \longrightarrow t = 13\]
\end{enumerate}

\item A store did \$40,000 in sales in 2016, and \$62,000 in 2018.
\begin{enumerate}[(a)]
\item Assuming the store's sales are growing linearly, find the growth rate $d$. \answersub{$d=11,000$}
\[62,000 = 40,000 + d(2) \longrightarrow 22,000 = 2d \longrightarrow d = 11,000\]
\item Write a linear model to describe this store's sales from 2016 onward. \answersub{$P_t = 40,000 + 11,000t$}
\item Predict the store's sales in 2025. \answersub{\$139,000}
\[P_t = 40,000 + 11,000(9) = 139,000\]
\item When do you expect the store's sales to exceed \$100,000? \answersub{$t=5.45$; about halfway through 2021}
\[100,000 = 40,000 + 11,000t \longrightarrow 60,000 = 11,000t \longrightarrow t = 5.45\]
\end{enumerate}

\item A small town had 340 homes in 2010, and by 2020, this had grown to 375.
\begin{enumerate}[(a)]
\item Assuming the number of homes is growing linearly, find the growth rate $d$. \answersub{$d=3.5$}
\[375 = 340 + d(10) \longrightarrow 35 = 10d \longrightarrow d = 3.5\]
\item Write a linear model to describe the number of homes in this town from 2010 onward. \answersub{$P_t = 340 + 3.5t$}
\item Predict how many homes there will be in this town in 2034. \answersub{424 homes}
\[P_t = 340 + 3.5(24) = 424\]
\item When do you expect the number of homes to exceed 500? \answersub{$t=45.7$; mid to late 2055}
\[500 = 340 + 3.5t \longrightarrow 160 = 3.5t \longrightarrow t = 45.7\]
\end{enumerate}

\item A population of beetles is growing according to a linear growth model.  Initially, there were 13 beetles, and 8 weeks later, there were 42 beetles.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of beetles over time, using weeks as the unit of time. \answersub{$P_t = 13 + 3.625t$}
\[42 = 13 + 8d \longrightarrow d = 3.625\]
\item How many beetles are there expected to be 14 weeks after the initial point? \answersub{approximately 64 beetles}
\[P_t = 13 + 3.625(14) = 63.75\]
\item When do you expect the number of beetles to exceed 70? \answersub{$t=15.7$}
\[70 = 13 + 3.625t \longrightarrow t = 15.7\]
\end{enumerate}

\item The number of streetlights in a town is growing linearly.  Four months ago there were 130 lights.  Now there are 146 lights.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of streetlights in the town over time, using months as the unit of time. \answersub{$P_t = 130 + 4t$}
\[146 = 130 + 4d \longrightarrow d = 4\]
\item How many streetlights are expected a year from \emph{now} (not a year from the beginning)? \answersub{194 streetlights}
\[P_t = 130 + 4(16) = 194\]
\item When do you expect the number of streetlights to exceed 200? \answersub{$t = 17.5$}
\[200 = 130 + 4t \longrightarrow t = 17.5\]
\end{enumerate}

\item In 1990, there were 112 nuclear power plants in the U.S.  By 2019, this number had fallen to 96.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of nuclear power plants from 1990 onward. \answersub{$P_t = 112 - 0.55t$}
\[96 = 112 + 29d \longrightarrow d = -0.55\]
\item Using this linear model, predict the number of nuclear power plants in the U.S. in 2030. \answersub{90 power plants}
\[P_t = 112 - 0.55(40) = 90\]
\item When do you expect the number of nuclear power plants to reach 80? \answersub{$t = 58.2$; early 2048}
\[80 = 112 - 0.55t \longrightarrow t = 58.2\] 
\end{enumerate}

\item In 1990, approximately 1,820,000 violent crimes were reported in the U.S.  By 2019, this number had fallen to approximately 1,200,000.
\begin{enumerate}[(a)]
\item Write a linear model to describe the number of violent crimes in the U.S. from 1990 onward. \answersub{$P_t = 1,820,000 - 21,379t$}
\[1,200,000 = 1,820,000 + 29d \longrightarrow d = -21,379\]
\item Using this linear model, predict the number of violent crimes in 2040. \answersub{751,050 violent crimes}
\[P_t = 1,820,000 - 21,379(50) = 751,050\]
\item When do you expect the number of violent crimes to reach 1,000,000? \answersub{$t = 38.4$; about halfway through 2028}
\[1,000,000 = 1,820,000 - 21,379t \longrightarrow t = 38.4\]
\end{enumerate}

\item The table below shows the average annual cost of health insurance for a single individual, from 1999 to 2019, according to the Kaiser Family Foundation.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Cost}\\
\hline
 & \\
1999 & \$2,196\\
2000 & \$2,471\\
2001 & \$2,689\\
2002 & \$3,083\\
2003 & \$3,383\\
2004 & \$3,695\\
2005 & \$4,024\\
2006 & \$4,242\\
2007 & \$4,479\\
2008 & \$4,704\\
2009 & \$4,824\\
2010 & \$5,049\\
2011 & \$5,429\\
2012 & \$5,615\\
2013 & \$5,884\\
2014 & \$6,025\\
2015 & \$6,251\\
2016 & \$6,435\\
2017 & \$6,690\\
2018 & \$6,896\\
2019 & \$7,186
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Using only the data from the first and last years, build a linear model to describe the cost of individual health insurance from 1999 onward. \answersub{$P_t = 2196 + 249.5t$}
\[7186 = 2196 + 20d \longrightarrow d = 249.5\]
\item Using this linear model, predict the cost of insurance in 2030. \answersub{\$9,930.50}
\[P_t = 2196 + 249.5(31) = 9930.5\]
\item According to this model, when do you expect the cost of individual insurance to reach \$12,000? \answersub{$t=39.3$; early 2038}
\[12,000 = 2196 + 249.5t \longrightarrow t = 39.3\]
\item Using a calculator or spreadsheet program, build a linear regression model to describe the cost of individual insurance from 1999 onward. \answersub{$P_t = 2381.7 + 243.97t$}
\item Using the regression model, predict the cost of insurance in 2030. \answersub{\$9,944.77}
\[P_t = 2381.7 + 243.97(31) = 9944.77\]
\item According to the regression model, when do you expect the cost of individual insurance to reach \$12,000? \answersub{$t=39.4$; early to mid 2038}
\[12,000 = 2381.7 + 243.97t \longrightarrow t = 39.4\]
\end{enumerate}

\item The table below shows the average annual cost of health insurance for a family, from 1999 to 2019, according to the Kaiser Family Foundation.
\begin{center}
\begin{tabular}{c c}
\textbf{Year} & \textbf{Cost}\\
\hline
 & \\
1999 & \$5,791\\
2000 & \$6,438\\
2001 & \$7,061\\
2002 & \$8,003\\
2003 & \$9,068\\
2004 & \$9,950\\
2005 & \$10,880\\
2006 & \$11,480\\
2007 & \$12,106\\
2008 & \$12,680\\
2009 & \$13,375\\
2010 & \$13,770\\
2011 & \$15,073\\
2012 & \$15,745\\
2013 & \$16,351\\
2014 & \$16,834\\
2015 & \$17,545\\
2016 & \$18,142\\
2017 & \$18,764\\
2018 & \$19,616\\
2019 & \$20,576
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
\item Using only the data from the first and last years, build a linear model to describe the cost of family health insurance from 1999 onward. \answersub{$P_t = 5,791 + 739.25t$}
\[20,576 = 5,791 + 20d \longrightarrow d = 739.25\]
\item Using this linear model, predict the cost of insurance in 2025. \answersub{\$25,011.50}
\[P_t = 5,791 + 739.25(26) = 25,011.5\]
\item According to this model, when do you expect the cost of family insurance to reach \$30,000? \answersub{$t=32.7$; mid to late 2031}
\[30,000 = 5,791 + 739.25t \longrightarrow t = 32.7\]
\item Using a calculator or spreadsheet program, build a linear regression model to describe the cost of family insurance from 1999 onward. \answersub{$P_t = 6050.3 + 724.72t$}
\item Using the regression model, predict the cost of insurance in 2025. \answersub{\$24,893.02}
\[P_t = 6050.3 + 724.72(26) = 24,893.02\]
\item According to the regression model, when do you expect the cost of family insurance to reach \$30,000? \answersub{$t = 33.0$; beginning of 2032}
\[30,000 = 6050.3 + 724.72t \longrightarrow t = 33.0\]
\end{enumerate}
\end{enumerate}